### 2019-10-17 scripts/02_renorm_and_other_cluster.Rmd

---
title: "renorm of the genefa and pathway abundance"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is to document what happened.

This step is done at the cluster

For the given valid samples("1725E" "2222A" "2245A" removed)
* merge them into one table (note that humann2 join table with take all of the tsv files that are in the folder)
* renorm to relab
* split them into stratified and unstratified

```{bash}
scp -r  daia1@lilac.mskcc.org:/home/daia1/my_workdir/samples/melody/humann2_final_out .
```