### 2019-10-17 scripts/04_lefse_ORR.Rmd

---
title: "lefse with ORR"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# because there are currently 36 valid ORR samples so it needs to be normalized separately.

* ORR & Toxicity

```{r}
# the two samples that are currenly NR in ORR
NRs <- read_csv('../data/current_pheno.csv') %>% 
  filter(ORR == "NR") %>% 
  pull(Sampleid)

# the current pathway abundance matrix that is unnormalized
abun <- read_tsv('../data/humann2_res/humann2_pathabundance.tsv') %>% 
  rename(sampleid = names(.)[1]) %>% 
  rename_all(funs(str_replace(., '_Abundance$',''))) %>% 
  select(-NRs)

ncol(abun)

# write out to do the normalization
abun %>% 
  write_tsv('../data/humann2_res/humann2_pathabundance_36.tsv')

```

```{bash}
# renorm
humann2_renorm_table --input humann2_pathabundance_36.tsv -u relab --output humann2_pathabundance_36_relab.tsv
```

```{bash}
# split the table
humann2_split_stratified_table --input humann2_pathabundance_36_relab.tsv --output humann2_final_out_36
```

```{r}
# load the split table with 36 and align the sampleid with the pheno table
cts <- read_tsv('../data/humann2_res/humann2_final_out_36/humann2_pathabundance_36_relab_unstratified.tsv')

# ORR  there is gonna be 36 samples cuz 2 are NR
pheno_O <- read_csv('../data/current_pheno.csv') %>% 
  filter(ORR != "NR") %>% 
  select(-CR, -Toxicity )

# reorder the cts table
cts <- bind_cols(cts[,1],cts[,pheno_O$Sampleid])

# check if the sampleids really align now
all.equal(colnames(cts)[2:ncol(cts)], pheno_O$Sampleid)

# write out the cts table for later binding
cts %>% 
  write_tsv('../data/cts_36_reorder.tsv')

```

# Different combination of the classes + subclasses

```{r}
# get different combination of the classes
classes <- c('ORR')
subclass <- c('Disease_Status_CAR','Disease_Class','CAR_Group')

combi <- expand.grid(classes = classes,subclass = subclass)

# just to be sure
all.equal(pheno_O$Sampleid, colnames(cts)[2:ncol(cts)])

# use each combi row in the above df to create the pheno table for the lefse input

apply(combi, 1, function(row) {
  pheno_O %>% 
    select(row[['classes']], row[['subclass']]) %>% 
    t() %>% 
    write.table(str_glue('../data/lefse_input/pheno__{row[["classes"]]}__{row[["subclass"]]}.txt'), sep = '\t', quote = F, row.names = T, col.names = F)
})
```

# Create the formatted input for lefse

```{r}
# create the commad in df with str_glue and then output to a bash script
pheno_combi <- data_frame(pheno_file = list.files('../data/lefse_input', pattern = 'pheno__ORR', full.names = T)) %>% 
  mutate(combi = str_extract(pheno_file, 'pheno.+\\.'))

pheno_combi %>% 
  mutate(cat_cmd = str_glue('cat {pheno_file} ../data/cts_36_reorder.tsv > ../data/lefse_input/pathway_abundace_{combi}tsv')) %>% 
  select(cat_cmd) %>% 
  write_csv('cat_combi_3.sh', col_names = F)
```

```{r}
# create the cmd to do the format input stuff
pheno_combi %>% 
  mutate(input_tsv = str_glue('../data/lefse_input/pathway_abundace_{combi}tsv'),
         fmt_input = str_replace(input_tsv, 'tsv','in'),
         fmt_cmd = str_glue('format_input.py {input_tsv} {fmt_input} -c 1 -s 2 -u 3 -o 1000000')) %>% 
  select(fmt_cmd) %>% 
  write_csv('format_input_3.sh', col_names = F)
```

```{r}  
# the actual step to run lefse
pheno_combi %>%  
  mutate(fmt_input = str_glue('../data/lefse_input/pathway_abundace_{combi}in'),
         lefse_res = str_replace(fmt_input, 'in$','res'),
         run_cmd = str_glue('run_lefse.py {fmt_input} {lefse_res}')) %>% 
  select(run_cmd) %>% 
  write_csv('run_lefse_3.sh', col_names = F)
```

```{r}
# plotting cmd 
pheno_combi %>% 
  mutate(lefse_res = str_glue('../data/lefse_input/pathway_abundace_{combi}res'),
         fea_png = str_glue('../figs/pathway_abundace_{combi}png'),
         cmd = str_glue('plot_res.py {lefse_res}  {fea_png} --feature_font_size 4 --width 10 --dpi 300 --title {combi}')) %>% 
  select(cmd) %>% 
  write_csv('plottint_de_feature_3.sh', col_names = F)

# No differentially abundant features found  
```