### 2019-10-17 scripts/09_barplot_myself.Rmd

---
title: "Humann2 bar plot by myself"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The stacked bar plot like humann2 did, but do it myself.

```{r}
# send the test file to google group
# read_tsv('../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.pcl', col_names = F) %>% 
#   filter(str_detect(X1, 'ARO-PWY|Toxicity'))  %>% 
#   write_tsv('../data/humann2_res/test_ARO-PWY_pw_abundance.tsv', col_names = F)
```

## Reorder the stratified pathabundance table and align with the pheno sampleid

```{r}
# pheno table 
pheno <- read_csv('../data/cleaned_pheno/current_pheno.csv')%>% 
  filter(Sampleid %in% valid_s$X1) %>% 
  select(-Censor_YN, -Censor_reason )

# the stratified table is used cuz we wanna know the species info
stratified <- read_tsv('../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified.tsv') %>% 
  rename(pathway = names(.)[1]) %>% 
  rename_all(funs(str_replace(., '_Abundance$','')))

stratified <- bind_cols(stratified[,1],stratified[,pheno$Sampleid])

all.equal(colnames(stratified)[2:ncol(stratified)], pheno$Sampleid)

# write out this reordered table for later use
stratified %>% 
  write_tsv('../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered.tsv')
```

### Starting from toxicity

```{r}
# get the pheno lable of the samples
phelabel <-  pheno %>% 
  select(Sampleid, Toxicity)

cnt <- stratified %>% 
  filter(str_detect(pathway, 'PWY-4242: pantothenate and coenzyme A biosynthesis III')) %>% 
  mutate(pathway = str_replace(pathway, 'PWY-4242: pantothenate and coenzyme A biosynthesis III\\|', ''))  %>% 
  # extract the species info
  mutate(pathway = str_extract(pathway, 's__.+$|unclassified'),
         pathway = str_replace(pathway, 's__','')) %>% 
  rename(species = pathway)

# get the top 7 taxa and the unclassfied and other for all the others 
# get the above table 
mean_s <- cnt %>% 
  column_to_rownames('species') %>% 
  as.matrix() %>% 
  apply(., 1, mean) %>% 
  sort(decreasing = T)

# top 8
n <- 8
toppw <- names(mean_s)[1:n]

# colors and dark gray and light gray
hex <- c(brewer.pal(n -1, name = "Dark2"), '#A9A9A9', '#696969')

# now reduce the abundance to 9 rows, the top 8 and "other"

# split on whether in 8 or in "other"
splitother <- cnt %>% 
  split(f = str_detect(.$species, paste0(toppw, collapse = '|')))

cleaned_cnt <- bind_rows(splitother %>% 
    pluck('TRUE'),
  splitother %>% 
    pluck('FALSE') %>% 
    gather(key = 'Sample', value = 'relab', names(.)[2]: names(.)[ncol(.)]) %>% 
    group_by(Sample) %>% 
    summarise(other = sum(relab)) %>% 
    spread(Sample , other) %>% 
    mutate(species = 'Other') %>% 
    select(species, pheno$Sampleid)) %>% 
  # rearrange the species order factor lever
  mutate(species = factor(species, levels = c(toppw[toppw != 'unclassified'] , 'Other', 'unclassified'))) %>% 
  # the long format is the one for plotting
  gather(key = 'Sample', value = 'relab', names(.)[2]: names(.)[ncol(.)]) %>% 
  left_join(phelabel %>% 
              rename(Sample = Sampleid), by = 'Sample') %>% 
  mutate(Toxicity = factor(Toxicity))

# to put different color to the background to the phenotype
# create a dataframe to show where the coloring start and end for each group
# how many phenotypes ?
num_pheno <- length(names(table(phelabel$Toxicity)))
# have pre-defined background color for the pheno (to better contrast them)
bg_col <- c('#00468B', '#EC0000')
# number of samples in different pheno 
# the first phenotype(appearing on the left most of the plot)
phe1_num <- phelabel %>% 
  mutate(Toxicity = factor(Toxicity)) %>% 
  filter(Toxicity == as.character(levels(.$Toxicity)[1])) %>% 
  nrow

background_col <- data_frame(
  col = bg_col,
  xstart = c(0, phe1_num) + 0.5,
  xend = c(phe1_num, nrow(phelabel))+ 0.5,
  toxicity = names(table(phelabel$Toxicity))
)

# the ggplot
ggplot() +
  geom_bar(data = cleaned_cnt, 
           aes(x = reorder(Sample, as.numeric(Toxicity)), y = relab, fill = species, color = Toxicity), stat = 'identity') +
  geom_rect(data = background_col, 
            aes(xmin = xstart, xmax = xend, ymin = -Inf, ymax = Inf), 
            fill = background_col$col, 
            alpha = 0.2) +
  geom_bar(data = cleaned_cnt, aes(x = reorder(Sample, as.numeric(Toxicity)), y = relab, fill = species), stat = 'identity') +
  #guides(alpha = FALSE) +
  scale_fill_manual(values = hex) +
  labs(y = 'relative abundance',
       x = 'Sample',
       title = str_glue('')) +
  theme_light() +
  theme(
    axis.text.x = element_blank(),
    panel.grid.major = element_blank(), 
    axis.ticks.x = element_blank(),  
    panel.grid.minor = element_blank()
    ) +
  ggsave('../figs/species_pw/PWY-4242: pantothenate and coenzyme A biosynthesis III.jpg', width = 12, height = 8, dpi = 300)
```

## Find how many sig pathways from lefse I wanna plot

```{r}
# load the table that has the unique sig pathways
sigpw <- read_csv('../data/unique_sig_pathway.csv')

# find how many species total it has in all of the lefse sig pathways
siglefse <- read_tsv('../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered.tsv') %>% 
  filter(str_detect(pathway,  paste0(pw$pwid, collapse = '|'))) %>% 
  separate(X1, into = c('pwid','taxa'), sep = '\\|')

sig_num <- siglefse %>% 
  split(f = .$pwid) %>% 
  map_dfc(~ nrow(.)) %>% 
  gather(key = 'pw', value = 'taxa_num') %>% 
  arrange(desc(taxa_num)) %>% 
  separate(pw, into = c('pwid','pw'), sep = ': ')

```