### 2019-10-17 scripts/09_barplot_myself.Rmd

---
title: "Humann2 bar plot by myself"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
library(tidyverse)
```

The stacked bar plot like humann2 did, but do it myself.

```{r}
# send the test file to google group
# read_tsv('../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.pcl', col_names = F) %>% 
#   filter(str_detect(X1, 'ARO-PWY|Toxicity'))  %>% 
#   write_tsv('../data/humann2_res/test_ARO-PWY_pw_abundance.tsv', col_names = F)
```

## Reorder the stratified pathabundance table and align with the pheno sampleid

```{r}
# pheno table 
pheno <- read_csv('../data/cleaned_pheno/current_pheno.csv')%>% 
  filter(Sampleid %in% valid_s$X1) %>% 
  select(-Censor_YN, -Censor_reason )

# the stratified table is used cuz we wanna know the species info
stratified <- read_tsv('../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified.tsv') %>% 
  rename(PATHWAYID = names(.)[1]) %>% 
  rename_all(funs(str_replace(., '_Abundance$','')))

stratified <- bind_cols(stratified[,1],stratified[,pheno$Sampleid])

all.equal(colnames(stratified)[2:ncol(stratified)], pheno$Sampleid)

# write out this reordered table for later use
stratified %>% 
  write_tsv('../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered.tsv')
```

## Find how many sig PATHWAYIDs from lefse I wanna plot

```{r}
# load the table that has the unique sig PATHWAYIDs
sigpw <- read_csv('../data/unique_sig_PATHWAYID.csv') %>% 
  # grab the pw id which is gonna be able to match the rownames in the stratified table
  mutate(pwid = str_extract(PATHWAYID, '^.+_PWY|^PWY_\\d{3,4}'))  %>% 
  mutate(pwid = str_replace_all(pwid, '_', '-'))

# find how many species total it has in all of the lefse sig PATHWAYIDs
siglefse <- stratified %>% 
  filter(str_detect(PATHWAYID,  paste0(sigpw$pwid, collapse = '|'))) %>% 
  separate(PATHWAYID, into = c('pwid','taxa'), sep = '\\|')

sig_num <- siglefse %>% 
  split(f = .$pwid) %>% 
  map_dfc(~ nrow(.)) %>% 
  gather(key = 'pw', value = 'taxa_num') %>% 
  arrange(desc(taxa_num)) %>% 
  separate(pw, into = c('pwid','pw'), sep = ': ', remove = F) %>% 
  # for plotting reason filer on >= 9 rows
  filter(taxa_num >= 9)

```

## Starting from toxicity

```{r}
species_compo_barplot <- function(PATHWAYID, PWFULLNAME) {
  # TWO INPUTS: 
  # The pheno variable: CR, Toxicity, ORR
  # The full PATHWAYID
  # get the pheno lable of the samples

  #PATHWAYID <- 'COA-PWY'
  phelabel <-  pheno %>% 
    select(Sampleid, Toxicity)

  cnt <- stratified %>% 
    filter(str_detect(pathway, PATHWAYID)) %>% 
    # extract the species info
    mutate(pathway = str_extract(pathway, 's__.+$|unclassified'),
           pathway = str_replace(pathway, 's__','')) %>% 
    rename(species = pathway)

  # get the top 7 taxa and the unclassfied and other for all the others 
  # get the above table 
  mean_s <- cnt %>% 
    gather(key = 'Sample', value = 'relab', names(.)[2]: names(.)[ncol(.)]) %>% 
    group_by(species) %>% 
    summarise(species_sum = sum(relab)) %>% 
    arrange(desc(species_sum)) %>% 
    pull(species)

  # top n (variable) including unclassified
  n <- 8
  toppw <- mean_s[1:n]

  # colors and dark gray and light gray
  hex <- c(brewer.pal(n -1, name = "Dark2"), '#A9A9A9', '#696969')

  # now reduce the abundance to 9 rows, the top 8 and "other" (the current will work if the df has >= 9 rows)

  # split on whether in 8 or in "other"
  splitother <- cnt %>% 
    split(f = str_detect(.$species, paste0(toppw, collapse = '|')))

  cleaned_cnt <- bind_rows(splitother %>% 
      pluck('TRUE'),
    splitother %>% 
      pluck('FALSE') %>% 
      gather(key = 'Sample', value = 'relab', names(.)[2]: names(.)[ncol(.)]) %>% 
      group_by(Sample) %>% 
      summarise(other = sum(relab)) %>% 
      spread(Sample , other) %>% 
      mutate(species = 'Other') %>% 
      select(species, pheno$Sampleid)) %>% 
    # rearrange the species order factor lever
    mutate(species = factor(species, levels = c(toppw[toppw != 'unclassified'] , 'Other', 'unclassified'))) %>% 
    # the long format is the one for plotting
    gather(key = 'Sample', value = 'relab', names(.)[2]: names(.)[ncol(.)])

  # get the sum of relab per sample cuz we wanna order by this
  sample_relab_sum <- cleaned_cnt %>% 
    group_by(Sample) %>% 
    summarise(sample_sum = sum(relab))

  final_cnt <- cleaned_cnt %>% 
    left_join(sample_relab_sum, by = 'Sample') %>% 
    left_join(phelabel %>% 
                rename(Sample = Sampleid), by = 'Sample') %>% 
    mutate(Toxicity = factor(Toxicity)) %>% 
    arrange(Toxicity, sample_sum)

  # to put different color to the background to the phenotype
  # create a dataframe to show where the coloring start and end for each group
  # how many phenotypes ?
  num_pheno <- length(names(table(phelabel$Toxicity)))
  # have pre-defined background color for the pheno (to better contrast them)
  bg_col <- c('#00468B', '#EC0000')
  # number of samples in different pheno 
  # the first phenotype(appearing on the left most of the plot)
  phe1_num <- phelabel %>% 
    mutate(Toxicity = factor(Toxicity)) %>% 
    filter(Toxicity == as.character(levels(.$Toxicity)[1])) %>% 
    nrow

  background_col <- data_frame(
    col = bg_col,
    xstart = c(0, phe1_num) + 0.5,
    xend = c(phe1_num, nrow(phelabel))+ 0.5
  )

  # the ggplot
  ggplot() +
    geom_bar(data = final_cnt, 
             aes(x = reorder(Sample, as.numeric(Toxicity)), y = relab, fill = species), stat = 'identity') +
    geom_rect(data = background_col, 
              aes(xmin = xstart, xmax = xend, ymin = -Inf, ymax = Inf), 
              fill = background_col$col, 
              alpha = 0.2) +
    geom_bar(data = final_cnt, aes(x = reorder(Sample, as.numeric(Toxicity)), y = relab, fill = species), stat = 'identity') +
    scale_fill_manual(values = hex ) +
    labs(y = 'relative abundance',
         x = 'Sample',
         title = str_glue('Response to Toxicity NO (blue background) and YES (red background):\nspecies composition of the pathway abundance\n{PATHWAYID}: {PWFULLNAME}')) +
    theme_light() +
    theme(
      axis.text.x =  element_text(angle = 40, hjust = 1, size = 8),
      panel.grid.major = element_blank(), 
      axis.ticks.x = element_blank(),  
      panel.grid.minor = element_blank()
      ) +
    ggsave(str_glue('../figs/species_pw/{PATHWAYID}.jpg'), width = 12, height = 8, dpi = 300)
}

map2(sig_num$pwid, sig_num$pw, ~ species_compo_barplot(.x, .y))

sig_num$pw
```