### 2019-10-18 scripts/10_barplot_genefamilies.Rmd
update on the genefamily lefse

---
title: "Humann2 bar plot for gene family"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

######################################## Bash ########################################

```{bash}
# needs to download the utility mapping db first
humann2_databases --download utility_mapping full /Users/daia1/melodyP/data/humann2_res
```

```{bash}
# regroup to KEGG Orthogroups (KOs)
humann2_regroup_table \
            --input ../data/humann2_res/humann2_genefamilies.tsv  \
            --groups uniref90_ko  \
            --output ../data/humann2_res/humann2_genefamilies_KO_regroup.tsv
```

```{bash}
# renorm to relab
humann2_renorm_table  --input ../data/humann2_res/humann2_genefamilies_KO_regroup.tsv -u relab --output ../data/humann2_res/humann2_genefamilies_KO_regroup_relab.tsv
```

```{bash}
# split into unstratified and ..
humann2_split_stratified_table --input ../data/humann2_res/humann2_genefamilies_KO_regroup_relab.tsv  --output ../data/humann2_res/genefamily_split_KO
```

######################################## R ########################################

## What features(gene orthologs) do you wanna see? Do lefse first.

```{r}

```