### 2019-10-20 scripts/12_lefse_genefam_term_ggplot.Rmd
haha able to retrieve the actual term of KO

---
title: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(KEGGREST)
```

## Load the lefse results

```{r}
# load the lefse res of tox
# the input : lefse res files
fns <- list.files('../data/lefse_genefam/', pattern = 'genefam_abundace_pheno__[CR|Toxicity].+res$', full.names = T)
fns_name <- list.files('../data/lefse_genefam/', pattern = 'genefam_abundace_pheno__[CR|Toxicity].+res$')

# join all of the tables feature together
feature <- fns %>% 
  set_names(fns_name) %>% 
  # when meeting parsing failures, you need to specify all column types.
  map(~ read_tsv(., col_names = c('genefam','xx','direction','score','pval'),  col_types = 'ccccc') %>% 
        filter(!is.na(score))) %>% 
  bind_rows(.id = 'group') %>% 
  mutate(group = str_replace(group, 'genefam_abundace_pheno__','')) %>% 
  mutate(group = str_replace(group, '.res$','')) %>% 
  select(-xx, -pval) %>% 
  mutate(KOID = str_replace(genefam, '\\.g.+$',''))

# change the "N" direction to be minus score
# feature <- bind_rows(
#   feature %>% 
#     split(.$direction) %>% 
#     pluck('N') %>% 
#     mutate(score = -score),
#   feature %>% 
#     split(.$direction) %>% 
#     pluck('Y') 
# ) %>% 
#   arrange(group, pathway, score)

# feature %>% 
#   write_csv('../data/lefse_genefam/with_unmapped_sig_feature.csv')

```

## Use KEGGREST pkg to get the actual term

### work on the Toxicity first

```{r}
tox_df <-  feature %>% 
  filter(group == 'Toxicity') %>% 
  split(cut_width(1:nrow(.), 10, boundary=0)) %>% 
  map(~ pull(.data = ., KOID) %>% 
        map(~ list(KO_name = keggGet(.)[[1]]$NAME,
             KO_def = keggGet(.)[[1]]$DEFINITION)) %>% 
        bind_rows()) %>% 
  bind_rows()

```