### 2019-10-21 scripts/13_PCoA_brayCurtis_KO_abundance.Rmd
PCoA with KO abundance

---
title: "PCoA with bray curtis on KO abundance"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(vegan)
```

Attempt on making the figure 7 A and B of paper https://www.nature.com/articles/srep34826
Making PCoA plot with Bray-Curtis distance on the KO abundace matrix

```{r}
# using the unstratified cts first
pheno <- read_csv('../data/current_pheno.csv') %>% 
  filter(! Sampleid %in% c("1725E", "2222A" ,"2245A"))

ko <- read_tsv('../data/humann2_res/genefamily_split_KO/humann2_genefamilies_KO_regroup_relab_unstratified.tsv') %>% 
  rename(KO = names(.)[1]) %>% 
  rename_all(funs(str_replace(., '_Abundance-RPKs$','')))

ko <- bind_cols(ko[,1],ko[,pheno$Sampleid]) 
all.equal(pheno$Sampleid, colnames(ko)[2:ncol(ko)])
```
```{r}

```