### 2019-10-21 scripts/13_PCoA_brayCurtis_KO_abundance.Rmd
no separation

---
title: "PCoA with bray curtis on KO abundance"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)  
library(tidyverse)
library(vegan)
```

Attempt on making the figure 7 A and B of paper https://www.nature.com/articles/srep34826
Making PCoA plot with Bray-Curtis distance on the KO abundace matrix

```{r}
# using the unstratified cts first
pheno <- read_csv('../data/current_pheno.csv') %>% 
  filter(! Sampleid %in% c("1725E", "2222A" ,"2245A"))

ko <- read_tsv('../data/humann2_res/genefamily_split_KO/humann2_genefamilies_KO_regroup_relab_unstratified.tsv') %>% 
  rename(KO = names(.)[1]) %>% 
  rename_all(funs(str_replace(., '_Abundance-RPKs$','')))

ko <- bind_cols(ko[,1],ko[,pheno$Sampleid]) 
all.equal(pheno$Sampleid, colnames(ko)[2:ncol(ko)])

# turn the ko table in to a matrix
ko_matrix <- ko %>% 
  column_to_rownames('KO') %>% 
  as.matrix()
```

```{r}
pw <- read_tsv('../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_unstratified.tsv') %>% 
  rename(pw = names(.)[1]) %>% 
  rename_all(funs(str_replace(., '_Abundance$','')))

pw <- bind_cols(pw[,1],pw[,pheno$Sampleid]) 
all.equal(pheno$Sampleid, colnames(pw)[2:ncol(pw)])

pw_matrix <- pw %>% 
  column_to_rownames('pw') %>% 
  as.matrix()
```

```{r}
# PCoA
pheno_col <- pheno %>% 
  mutate(tox_col = if_else(Toxicity == 'Y', 'red', 'green '),
         CR_col = if_else(CR == 'Y', 'red', 'green '))

bc_dist <-  vegdist(t(pw_matrix), "bray")
pc_bc <- cmdscale(bc_dist, k = 2)
plot(pc_bc[,1], pc_bc[,2], col = pheno_col$tox_col, main = 'Pathway abundance color by tox')
plot(pc_bc[,1], pc_bc[,2], col = pheno_col$CR_col, main = 'Pathway abundance color by CR')
```

```{r}
# do the PcoA with bray curtis distance  
bc_dist <-  vegdist(t(ko_matrix), "bray")
pc_bc <- cmdscale(bc_dist, k = 2)
plot(pc_bc[,1], pc_bc[,2], col = pheno_col$tox_col, main = 'KO abundance color by tox')
plot(pc_bc[,1], pc_bc[,2], col = pheno_col$CR_col, main = 'KO abundance color by CR')
```

It just doesn't separate whether using gene ortholog or pathway abundance.