### 2019-10-21 scripts/14_pathway_KO_regroup.Rmd

---
title: "pathway KO regroup"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

```{bash}
# regroup to KEGG Orthogroups (KOs)
humann2_regroup_table \
            --input ../data/humann2_res/humann2_pathabundance.tsv  \
            --groups uniref90_pfam  \
            --output ../data/humann2_res/humann2_pathabundance_pfam_regroup.tsv
```

```{bash}
# renorm to relab
<!-- humann2_renorm_table  --input ../data/humann2_res/humann2_genefamilies_KO_regroup.tsv -u relab --output ../data/humann2_res/humann2_genefamilies_KO_regroup_relab.tsv -->
```

```{bash}
# split into unstratified and ..
<!-- humann2_split_stratified_table --input ../data/humann2_res/humann2_genefamilies_KO_regroup_relab.tsv  --output ../data/humann2_res/genefamily_split_KO -->
```