### 2019-10-24 scripts/01_clean_pheno.Rmd
update

---
title: "I mean really clean the pheno and get to know the hell of the samples"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
library(tidyverse)
library(readxl)
library(RVenn)
```

## Valid patient samples

```{r}
# the newest pheno table from M
# this sheet have nice sampleids for the samples, and also this is the table has all of the info
phe_M <- read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/CTC Event Table 10222019.xlsx',
                    sheet = 5)
```

```{r}
# selecting the columns relavant to this study
phe_M_clean <- phe_M %>% 
   rename_all(funs(str_replace_all(., ' \\(.+$','') %>% 
    str_replace_all(., ' ','_') %>% 
    str_replace(., ':','_'))) %>% 
  # select the columns that may be relavant
  select(Sampleid = Baseline_Sample_ID, 
         Censor_reason,
         CR, 
         ORR, 
         Toxicity, 
         Disease_Status_CAR = Key__Patient_Disease_Status_at_CAR, 
         Disease_Class, 
         CAR_Group) %>% 
  filter(is.na(Censor_reason))

nrow(phe_M_clean)

head(phe_M_clean)

# output this pheno table cuz its gonna be used downstream
phe_M_clean %>% 
  write_csv('../data/cleaned_pheno/current_pheno.csv')
```

## Valid shotgun data

```{r}
# loading the table from John to see how many samples we actually have shotgun data
have_sg <- read_csv('../data/baselinesforAngel10032019.csv') %>% 
  filter(shotgun_YN == 'yes') %>% 
  # removig 1943A cuz its low quality
  filter(sample_id != '1943A') %>% 
  # 171 is back, removing the the other two that doesn't make to the final valid sample
  filter(!sample_id %in% c('161','173'))

nrow(have_sg)

```

## Overlap: the real valid shotgun samples

```{r}
# visualize the difference 
two <- Venn(list(
  have_file = have_sg$sample_id,
  useable = phe_M_clean$Sampleid
))

ggvenn(two)      
```

```{r}
# write out the overlap sampleid cuz that's gonna be the sample_names in use on the cluster for normalization
intersect(have_sg$sample_id, phe_M_clean$Sampleid) %>% 
  write.table('../data/input_for_cluster_sampleid/valid_sample_names.txt', col.names = F, row.names = F, quote = F)
```