### 2019-10-31 scripts/16_RF_pathway.Rmd

---
title: "RF pathway abundance"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(randomForest)
library(ggpubr)
library(tidyverse)
library(kableExtra)
```

```{r}
# load the pheno
pheno <- read_csv('../data/cleaned_pheno/final_pheno.csv') %>% 
  select(-CAR_Group, -Censor_reason ) %>% 
  mutate_at(vars(str_which(names(.),'Sampleid', negate = T)), factor)

# the pathway abundance table
pabun <- read_tsv('../data/humann2_res/humann2_final_out/humann2_pathabundance_cpm_unstratified.tsv') %>% 
  rename(pathway = names(.)[1]) %>% 
  rename_all(funs(str_replace(., '_Abundance$',''))) %>% 
  # remove the unmapped and the unintegrated record
  filter(!pathway %in% c('UNMAPPED','UNINTEGRATED'))

pabun <- bind_cols(pabun[,1],pabun[,pheno$Sampleid])

pabun_rf <- pabun %>% 
  column_to_rownames('pathway') %>% 
  as.matrix() %>% 
  t

all.equal(rownames(pabun_rf), pheno$Sampleid)

tox_RF_PW <- randomForest(pabun_rf, factor(pheno$Toxicity), importance = T)

impor_pw <- importance(tox_RF_PW, type=1) %>% 
  as.data.frame() %>% 
  rownames_to_column(var = 'pathway') %>% 
  arrange(MeanDecreaseAccuracy) 
```