### 2019-11-05 scripts/19_phylum_table.Rmd
the correlation

---
title: "Look at the phylum level"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggpubr)
```

```{r}
phy <- read_tsv('../data/metaphlan2_result/merged_abundance_table_phylum.txt') %>% 
  rename_all(funs(str_replace(., '_profile$','')))

pheno <- read_csv('../data/cleaned_pheno/final_pheno.csv') %>% 
  select(-CAR_Group, -Censor_reason )

#mutate_at(vars(str_which(names(.),'Sampleid', negate = T)), factor)

phy <- bind_cols(phy[,1],phy[,pheno$Sampleid])

all.equal(colnames(phy)[2:ncol(phy)], pheno$Sampleid)

```

```{r}
phy %>% 
  gather(key = 'Sampleid', value = 'relab', names(.)[2]: names(.)[ncol(.)]) %>% 
  left_join(pheno, by = 'Sampleid') %>% 
  ggboxplot(x = 'ID', y = 'relab', color = 'Toxicity', palette  = 'lancet', add = 'jitter')  +
  theme(axis.text.x = element_text(angle = 30, hjust = 1)) 
```

```{r}
phy %>% 
  gather(key = 'Sampleid', value = 'relab', names(.)[2]: names(.)[ncol(.)]) %>% 
  left_join(pheno, by = 'Sampleid') %>% 
  filter(ID == 'Bacteroidetes') %>% 
  select(ID, relab, Toxicity) %>% 
  ggboxplot(x = 'Toxicity', y = 'relab', color = 'Toxicity', palette  = 'lancet', add = 'jitter')   + 
  stat_compare_means()+
  ggsave('../figs/phylum_abundance_Bacteroidetes.jpg')
```

```{r}
phy %>% 
  gather(key = 'Sampleid', value = 'relab', names(.)[2]: names(.)[ncol(.)]) %>% 
  left_join(pheno, by = 'Sampleid') %>% 
  filter(ID == 'Firmicutes') %>% 
  select(ID, relab, Toxicity) %>% 
  ggboxplot(x = 'Toxicity', y = 'relab', color = 'Toxicity', palette  = 'lancet', add = 'jitter')   + 
  stat_compare_means()  +
  ggsave('../figs/phylum_abundance_Firmicutes.jpg')
```