### 2019-11-05 scripts/20_correlation_phylum_bw_16S_shotgun.Rmd
update

---
title: "Correlation between 16S and shotgun"
author: "Anqi Dai"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
library(tidyverse)
library(ggpubr)
library(RColorBrewer)
library(kableExtra)
```

## Load the two datasets

```{r}
# shotgun
gun <- read_tsv('../data/metaphlan2_result/merged_abundance_table_phylum.txt') %>% 
  rename_all(funs(str_replace(., '_profile$',''))) %>% 
  rename(phylum = ID)

pheno <- read_csv('../data/cleaned_pheno/final_pheno.csv') %>% 
  select(-CAR_Group, -Censor_reason )
```

```{r}
dada2 <- read_rds('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/data/asv_set.RDS')

# how many samples are in the dada2 data
dada2_nsamps <- table(dada2 %>% 
  pluck('asv_counts') %>% 
    pull(sampleid)) %>% 
  names %>% 
  length

# get the relab at phylum level
sixteen <- dada2 %>% 
  pluck('asv_counts') %>% 
  select(ASV, count_relative, sampleid) %>% 
  left_join(dada2 %>% 
    pluck('asv_annotation') %>% 
    select(ASV, phylum ), by = 'ASV') %>% 
  filter(!is.na(phylum)) %>% 
  # select the 33 that are in the shotgun
  filter(sampleid %in% pheno$Sampleid) %>% 
  group_by(sampleid, phylum) %>% 
  summarise(relab = sum(count_relative)) %>% 
  spread(key = 'sampleid', value = 'relab', fill = 0)

```

```{r}
# same phylum term
ol_phylum <- intersect(sixteen$phylum, gun$phylum)

# have an idea that i better put the two of them table in the same table with overlapped phylum terms

two <- list(shotgun = gun,
            s16S = sixteen) %>% 
  imap(~  ..1 %>% 
      filter(phylum %in% ol_phylum) %>% 
      gather(key = 'sampleid', value = 'relab', names(.)[2] : names(.)[ncol(.)]) %>% 
      mutate(Group = ..2) %>% 
      mutate_at(vars(str_which(names(.),'relab', negate = T)), factor)
  )

Two <- two %>% 
  pluck('shotgun') %>% 
  rename(shotgun_relab = relab) %>% 
  select(-Group) %>% 
    full_join(two %>% 
    pluck('s16S') %>% 
    rename(s16S_relab = relab) %>% 
    select(-Group)) %>%  
  # to uniform the two data
  mutate(s16S_relab = s16S_relab*100)

```

## Compare with Spearman correlation for each sample

```{r}
Two %>% 
  ggplot(aes(x = shotgun_relab, y = s16S_relab, col = phylum)) +
    geom_point() +
    facet_wrap(~sampleid) +
  scale_color_manual(values = brewer.pal(n = length(ol_phylum), name = "Dark2")) +
  labs(x = 'Shotgun data',
       y = '16S data',
       title = 'Relative abundance comparison between shotgun and 16S data\nat phylum level') +
  ggsave('../figs/correlation/cor_between_16_and_shotgun.jpg', width = 10, height = 10, dpi = 300)

spearman_cor <- Two %>% 
  split(.$sampleid) %>% 
  map_dfr(~ round(cor.test(.$shotgun_relab, 
                           .$s16S_relab, 
                           method = 'spearman', 
                           exact = F)$estimate, 2)) %>% 
  gather(key = 'sampleid', value = 'rho') %>% 
  arrange(rho)

spearman_cor %>% 
  kable(caption = 'Spearman correlation at phylum level') %>% 
  kable_styling(full_width = T, position = 'left')

spearman_cor %>% 
  write_csv('../data/spearman_cor_phylum.csv')
```