### 2019-11-12 scripts/22_shortbred.Rmd
update

---
title: "Shortbred results"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(kableExtra)
library(ggpubr)
```

```{r}
# create symlinks for the fastq files to work in the quant step
pheno <- read_csv('../data/cleaned_pheno/final_pheno.csv')  %>% 
  select(Sampleid)  %>% 
  mutate(fastq = str_glue('{Sampleid}.fastq'),
         idx = seq(1, nrow(.)),
         idx_fn = str_glue('WGS_{idx}.fastq'),
         cmd = str_glue('ln -s /home/daia1/my_workdir/samples/melody/{fastq} /home/daia1/my_workdir/samples/melody/{idx_fn}'))

pheno %>% 
  select(cmd) %>% 
  write_tsv('../scripts/link_wgs_fastq.sh', col_names = F)
```