### 2019-11-14 scripts/23_unifrac.Rmd
update

---
title: "Unifrac tree"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(seqinr)
source('~/MSK/work/microbiome_db/SQL/scripts/get_data_from_query_OTU.R')
```

```{r}
# load the table
annot <- get_data_from_query_OTU(0,'asv_annotation_blast_ag')  
seqs <- get_data_from_query_OTU(0,'asv_sequences_ag') 
```

```{r}
# create fasta file and write out to disk which will do the msa later
res <- apply(seqs, 1, function(Row){
  write.fasta(sequences = as.list(Row[['asv_sequence']]), 
              names = Row[['key']],
              file.out = str_glue('../output/ASV_fa/{Row[["key"]]}_ASV_seq.fa'))
})

# should be a fasta file with 80,000 sequences 
```

```{bash}
# do the msa
```