### 2019-11-26 scripts/24_pathcoverage.Rmd

---
title: "pathway coverage"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

```{r}
pwc <- read_tsv('~/melodyP/data/humann2_res/humann2_final_out/humann2_pathcoverage_unstratified.tsv')
```