### 2019-11-27 scripts/11_lefse_geneorthologs.Rmd

---
title: "Lefse on gene orthologs"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Make pheno table(with only classes)

```{r}
# loading the table with the valid sampleids
pheno <- read_csv('../data/cleaned_pheno/final_pheno.csv')

```

```{r}
# get different combination of the classes
classes <- c('CR', 'Toxicity')

#subclass <- c('Disease_Status_CAR','Disease_Class','CAR_Group')
combi <- expand.grid(classes = classes)

# use each combi row in the above df to create the pheno table for the lefse input
apply(combi, 1, function(row) {
  pheno %>% 
    mutate(CR = str_glue('CR_{CR}'),
           Toxicity = str_glue('Toxicity_{Toxicity}')) %>% 
    select(row[['classes']]) %>% 
    t() %>% 
    write.table(str_glue('../data/lefse_genefam/pheno__{row[["classes"]]}.txt'), sep = '\t', quote = F, row.names = T, col.names = F)
})
```

## Reorder the KO stratified table

```{bash}
# get the KO regrouped table
humann2_regroup_table \
            --input ../data/humann2_res/humann2_genefamilies.tsv  \
            --groups uniref90_ko  \
            --output ../data/humann2_res/humann2_genefamilies_KO_regroup.tsv
```

```{r}
stra_go <- read_tsv('../data/humann2_res/genefamily_split_KO/humann2_genefamilies_KO_regroup_relab_stratified.tsv') %>% 
  rename(sampleid = names(.)[1]) %>% 
  rename_all(funs(str_replace(., '_Abundance-RPKs$',''))) %>% 
  # remove the ungrouped since it's not really giving gene fam info
  filter(! str_detect(sampleid, '^UNGROUPED'))

stra_go <- bind_cols(stra_go[,1],stra_go[,pheno$Sampleid])

all.equal(pheno$Sampleid, colnames(stra_go)[2:ncol(stra_go)])

# write out

stra_go %>% 
  write_tsv('../data/humann2_res/genefamily_split_KO/humann2_genefamilies_KO_regroup_relab_stratified_reordered.tsv')
```

## str_glue

```{r}
# create the commad in df with str_glue and then output to a bash script
pheno_combi <- data_frame(pheno_file = list.files('../data/lefse_genefam', pattern = '.txt$', full.names = T)) %>% 
  mutate(combi = str_extract(pheno_file, 'pheno.+\\.'))

pheno_combi %>% 
  mutate(cat_cmd = str_glue('cat {pheno_file} ../data/humann2_res/genefamily_split_KO/humann2_genefamilies_KO_regroup_relab_stratified_reordered.tsv > ../data/lefse_genefam/genefam_abundace_{combi}tsv')) %>% 
  select(cat_cmd) %>% 
  write_csv('cat_combi_genef.sh', col_names = F)
```

```{r}
# create the cmd to do the format input stuff
pheno_combi %>% 
  mutate(input_tsv = str_glue('../data/lefse_genefam/genefam_abundace_{combi}tsv'),
         fmt_input = str_replace(input_tsv, 'tsv','in'),
         fmt_cmd = str_glue('format_input.py {input_tsv} {fmt_input} -c 1 -u 2 -o 1000000')) %>% 
  select(fmt_cmd) %>% 
  write_csv('format_input_genef.sh', col_names = F)
```

```{r}
# the actual step to run lefse
pheno_combi %>% 
  mutate(fmt_input = str_glue('../data/lefse_genefam/genefam_abundace_{combi}in'),
         lefse_res = str_replace(fmt_input, 'in$','res'),
         run_cmd = str_glue('run_lefse.py {fmt_input} {lefse_res}')) %>% 
  select(run_cmd) %>% 
  write_csv('run_lefse_genef.sh', col_names = F)
```

```{r}
# plotting cmd 
pheno_combi %>% 
  mutate(lefse_res = str_glue('../data/lefse_genefam/genefam_abundace_{combi}res'),
         fea_png = str_glue('../figs/genefam_abundace_{combi}png'),
         cmd = str_glue('plot_res.py {lefse_res}  {fea_png} --feature_font_size 4 --width 10 --dpi 300 --title {combi}')) %>% 
  select(cmd) %>% 
  write_csv('plotting_de_feature_genef.sh', col_names = F)
```

## bash

```{bash}
bash cat_combi_genef.sh
bash format_input_genef.sh
bash run_lefse_genef.sh

```