### 2019-11-27 scripts/25_followups.Rmd

---
title: "Two patient follow up"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

```{r}
target <- read_csv('/Volumes/vandenBrinkLab/Angel_Dai/Full_human_shotgun_catalog/full_human_shotgun_catalog_updated.csv') %>% 
  # the patient 1487 and 2050
  filter(str_detect(sampleid, '1487|2050')) %>% 
  filter(!str_detect(directory, 'A_IGO'))

target %>% 
  dplyr::select(directory) %>% 
  mutate(cmd = str_glue('rsync --progress --partial -avz {directory} daia1@lilac.mskcc.org:~/my_workdir/samples/melody')) %>% 
  dplyr::select(cmd) %>%   
  write_csv('../data/upload_M_5.sh', col_names = F)
```