### 2019-12-03 scripts/26_unpack_pw.Rmd
update

---
title: "Find gene fam contribution to pw"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

```{bash}
# renorm the gene family table
humann2_renorm_table  --input humann2_genefamilies.tsv -u cpm -s n --output humann2_genefamilies_cpm.tsv

# unpack the pw using humann2 script
humann2_unpack_pathways --input-genes humann2_genefamilies_cpm.tsv --input-pathways humann2_pathabundance_cpm.tsv \
  --remove-taxonomy --output humann2_pwa_unpacked.tsv 
```

```{r}
# look at the results
unpacked <- read_tsv('../data/humann2_res/humann2_pwa_unpacked.tsv', skip = 1, col_names = c('pw','cpm'))

# the pantothenate (PWY-4242)
vb5 <- unpacked %>% 
  filter(str_detect(pw, '4242'))
```