### 2019-12-03 scripts/26_unpack_pw.Rmd
unpacked

---
title: "Find gene fam contribution to pw"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggpubr)
```

```{bash}
# renorm the gene family table
humann2_renorm_table  --input humann2_genefamilies.tsv -u cpm -s n --output humann2_genefamilies_cpm.tsv

# unpack the pw using humann2 script
humann2_unpack_pathways --input-genes humann2_genefamilies_cpm.tsv --input-pathways humann2_pathabundance_cpm.tsv \
  --remove-taxonomy --output humann2_pwa_unpacked.tsv 
```

```{r}
# look at the results
unpacked <- read_tsv('../data/humann2_res/humann2_pwa_unpacked.tsv', skip = 1, col_names = c('pw','cpm'))

# the pantothenate (PWY-4242)

# get the top N results
N <- 20

vb5 <- unpacked %>% 
  filter(str_detect(pw, '4242')) %>% 
  filter(str_detect(pw, '\\|')) %>% 
  arrange(desc(cpm)) %>% 
  separate(pw, into = c('name','uniref90_ID'), '\\|') %>% 
  top_n(N)

vb5 %>% 
  ggbarplot(x = 'uniref90_ID', y = 'cpm', color = 'white', fill = 'forestgreen', title = 'PWY-4242')  +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggsave('../figs/unpack/PWY-4242.jpg', width = 8, height = 7, dpi = 300)

# COA 
unpacked %>% 
  filter(str_detect(pw, 'COA-PWY\\|')) %>% 
  filter(str_detect(pw, '\\|')) %>% 
  arrange(desc(cpm)) %>% 
  separate(pw, into = c('name','uniref90_ID'), '\\|') %>% 
  top_n(N) %>% 
  ggbarplot(x = 'uniref90_ID', y = 'cpm', color = 'white', fill = 'forestgreen', title = 'COA-PWY')  +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggsave('../figs/unpack/COA-PWY.jpg', width = 8, height = 7, dpi = 300)

# PANTOSYN-PWY
unpacked %>% 
  filter(str_detect(pw, 'PANTOSYN-PWY\\|')) %>% 
  filter(str_detect(pw, '\\|')) %>% 
  arrange(desc(cpm)) %>% 
  separate(pw, into = c('name','uniref90_ID'), '\\|') %>% 
  top_n(N) %>% 
  ggbarplot(x = 'uniref90_ID', y = 'cpm', color = 'white', fill = 'forestgreen', title = 'PANTOSYN-PWY')  +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggsave('../figs/unpack/PANTOSYN-PWY.jpg', width = 8, height = 7, dpi = 300)
```