### 2020-01-08 scripts/30_univariate_lm.Rmd
update

---
title: "Univariate lm model"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

```{r}
final_pheno <- read_csv('../data/cleaned_pheno/final_pheno.csv')

pw <- read_tsv('../data/humann2_res/humann2_final_out/humann2_pathabundance_cpm_unstratified.tsv') %>% 
  dplyr::rename_all(funs(str_replace(., '_Abundance$',''))) %>% 
  dplyr::rename(pw = names(.)[1])  %>% 
  # remove the things after the : so that its easier to see on the plot(the pathway name)
  mutate(pw = str_replace(pw, ':.+$',''))

pw <- bind_cols(pw[,1],pw[,final_pheno$sampleid])

all.equal(colnames(pw)[2:ncol(pw)], final_pheno$sampleid)
```

```{r}
# filter to have each pathway with at least 30% of nonzero (zero <= 50%)
pw_fil <- pw %>% 
  column_to_rownames('pw') %>% 
  as.matrix()

num_zero_thre <- floor(ncol(pw_fil) * 0.5)

pw_fil <- pw_fil[rowSums(pw_fil == 0) <= num_zero_thre, ]
nrow(pw_fil)

```

```{r}
# select the most abundant N pathways

N <- 10
topN_pw <- sort(rowSums(pw_fil), decreasing = T)[1:N]
pw_Fil <- pw_fil[names(topN_pw),]

all.equal(colnames(pw_Fil), final_pheno$sampleid)

# do a linear regression of each rows and then do a FDR correction later
# the pheno in the same order as the current sample names 
phe_tox <- final_pheno$Toxicity
phe_CR <- final_pheno$CR

res_tox <- apply(pw_Fil, 1, function(r){
  fit = summary(lm(r ~ phe_tox))$coefficients
  return(pval = fit[nrow(fit),ncol(fit)])
}) %>% 
  data.frame() %>% 
  rename(pval = names(.)[1]) %>% 
  rownames_to_column('pw') %>% 
  mutate(padj = p.adjust(pval, method = 'BH')) %>% 
  arrange(desc(padj))

res_CR <- apply(pw_Fil, 1, function(r){
  fit = summary(lm(r ~ phe_CR))$coefficients
  return(pval = fit[nrow(fit),ncol(fit)])
}) %>% 
  data.frame() %>% 
  rename(pval = names(.)[1]) %>% 
  rownames_to_column('pw') %>% 
  mutate(padj = p.adjust(pval, method = 'BH')) %>% 
  arrange(desc(padj))
```