### 2020-01-16 scripts/32_two_genera.Rmd
update

---
title: "Two genera pathway association"
author: "Anqi Dai"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
library(tidyverse)
library(pheatmap)
library(kableExtra)
```

Load the temporary table of humann2 gene fam in genus level results

Renorm

Split

# The genus of Enterococcus

Get the Enterococcus count of the pathway

```{r}
pheno <- read_csv('../data/cleaned_pheno/final_pheno.csv')

genus <- read_tsv('../data/humann2_res/genus_level/humann2_pathabundance_genus_level_cpm_stratified.tsv') %>% 
  dplyr::rename(pw = names(.)[1]) %>% 
  rename_all(funs(str_replace(., '_genefamilies_.+$','')))

genus_entero <- genus %>% 
  filter(str_detect(pw, 'g__Enterococcus')) 
```

```{r}
# a heatmap to see the counts distribution

annot <- data.frame(row.names = pheno$sampleid, tox = pheno$Toxicity, CR = pheno$CR)

pheatmap(
  genus_entero %>% 
    column_to_rownames('pw') %>% 
    select(-`1409B`) %>% 
    as.matrix,
  #color=bwrPalette, 
  annotation_col = annot,
  show_rownames = TRUE,
  show_colnames = TRUE,
  filename = '../figs/Enterococcus_heatmap.jpg',
  height = 18,
  width = 18,
  cluster_rows = F,
  cluster_cols = F
)
```

Many of them are zeros that we can see, I don't think it makes a lot of sense to do a linear regression.

## Do a wilcox in relation to Toxicity

```{r}
# ehh do it anyway
entero_matrix <- genus_entero %>% 
    column_to_rownames('pw') %>% 
    as.matrix

# rearrange the column names
entero_matrix <- entero_matrix[,pheno$sampleid]

all.equal(colnames(entero_matrix), pheno$sampleid)

# a linear regression of CR and tox respectively
sig_tox <- apply(entero_matrix, 1, function(r){
    res = wilcox.test(r ~ pheno$Toxicity, exact = F)
    return(pval = res$p.value)
  }) %>% 
    data.frame() %>% 
    dplyr::rename(pval = names(.)[1]) %>% 
    rownames_to_column('pw') %>% 
    mutate(padj = p.adjust(pval, method = 'BH')) %>% 
    arrange(padj)

sig_tox %>% 
  head() %>% 
  kable() %>% 
  kable_styling(full_width = F)

sig_CR <- apply(entero_matrix, 1, function(r){
    res = wilcox.test(r ~ pheno$CR, exact = F)
    return(pval = res$p.value)
  }) %>% 
    data.frame() %>% 
    dplyr::rename(pval = names(.)[1]) %>% 
    rownames_to_column('pw') %>% 
    mutate(padj = p.adjust(pval, method = 'BH')) %>% 
    arrange(padj)

sig_CR %>% 
  head() %>% 
  kable() %>% 
  kable_styling(full_width = F)
```

# Gather all the genera in Bacteroidetes

```{r}
bac_genera_g <- c('Bacteroides',
                'Prevotella',
                'Parabacteroides',
                'Porphyromonas',
                'Flavobacterium',
                'Cytophaga',
                'Alistipes',
                'Flexibacter',
                'Capnocytophaga',
                'Dyadobacter',
                'Hymenobacter',
                'Barnesiella',
                'Dorea',
                'Flectobacillus',
                'Arcicella',
                'Chryseobacterium',
                'Salinibacter',
                'Polaribacter',
                'Cellulophaga',
                'Blattabacterium',
                'Runella',
                'Flammeovirga',
                'Tenacibaculum',
                'Sphingobacterium',
                'Pedobacter',
                'Leeuwenhoekiella',
                'Butyricimonas',
                'Bergeyella',
                'Robiginitalea',
                'Croceibacter',
                'Dysgonomonas',
                'Algoriphagus',
                'Thermonema',
                'Gelidibacter',
                'Arenibacter',
                'Alkaliflexus',
                'Aequorivita',
                'Muricauda',
                'Chitinophaga',
                'Epilithonimonas',
                'Paludibacter',
                'Belliella',
                'Butyricicoccus',
                'Megamonas',
                'Lewinella',
                'Niabella',
                'Vitellibacter',
                'Fabibacter',
                'Pontibacter',
                'Proteiniphilum',
                'Mucilaginibacter')

bac_genera <- read_csv('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/data/Bacteroidetes_genus.csv') %>% 
  pull(genus)

intersect(bac_genera, bac_genera_g)
```

```{r}
# find out all the rows that have bac genera
genus_bact <- genus %>% 
  separate(pw, into = c('pathway','genus'), '\\|') %>% 
  mutate(genus = str_replace(genus, 'g__','')) %>% 
  filter(genus %in% bac_genera)

genus_bact  %>% 
  count(pathway)

# group by the pathway and get the summed count of all the bact genera
bact_matrix <- genus_bact %>% 
  select(-genus) %>% 
  gather(key = 'sampleid', value = 'CPM', names(.)[2]:names(.)[ncol(.)]) %>% 
  group_by(pathway, sampleid) %>% 
  summarise(cpm = sum(CPM)) %>% 
  spread(key = 'sampleid', value = 'cpm') %>% 
  column_to_rownames('pathway') %>% 
  as.matrix

bact_matrix <- bact_matrix[,pheno$sampleid]

all.equal(colnames(bact_matrix), pheno$sampleid)
```
## a wilcox of CR and tox respectively

```{r}
# a wilcox of CR and tox respectively
sig_tox <- apply(bact_matrix, 1, function(r){
    res = wilcox.test(r ~ pheno$Toxicity, exact = F)
    return(pval = res$p.value)
  }) %>% 
    data.frame() %>% 
    dplyr::rename(pval = names(.)[1]) %>% 
    rownames_to_column('pw') %>% 
    mutate(padj = p.adjust(pval, method = 'BH')) %>% 
    arrange(padj)

sig_tox %>% 
  head() %>% 
  kable() %>% 
  kable_styling(full_width = F)

sig_CR <- apply(bact_matrix, 1, function(r){
    res = wilcox.test(r ~ pheno$CR, exact = F)
    return(pval = res$p.value)
  }) %>% 
    data.frame() %>% 
    dplyr::rename(pval = names(.)[1]) %>% 
    rownames_to_column('pw') %>% 
    mutate(padj = p.adjust(pval, method = 'BH')) %>% 
    arrange(padj)

sig_CR %>% 
  head() %>% 
  kable() %>% 
  kable_styling(full_width = F)
```