### 2020-01-19 scripts/34_secondary_bile.Rmd
update

---
title: "secondary bile acid biosynthesis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggpubr)
```

```{r}
pheno <- read_csv('../../CART_microbiome/data/cleaned_pheno/final_pheno.csv')
mbile <- list.files('../../CART_microbiome/data/bile/', full.names = T)

all <- mbile %>% 
  set_names(mbile) %>% 
  map(~ read_tsv(.)) %>% 
  bind_rows(.id = 'sampleid') %>% 
  group_by(sampleid) %>% 
  summarise(RPKM = sum(Count))

all %>% 
  mutate(sampleid = str_replace(sampleid, '.+//',''),
         sampleid = str_replace(sampleid, '_short.+$','')) %>% 
  full_join(pheno, by = 'sampleid') %>% 
  ggboxplot(x = 'CR', y = 'RPKM', color = 'CR', palette = 'lancet', add = 'dotplot')  +
  stat_compare_means(comparisons = list(c('Y', 'N')), 
                     label = "p.signif",
                     method = 'wilcox.test',
                     correct=FALSE)

all %>% 
  mutate(sampleid = str_replace(sampleid, '.+//',''),
         sampleid = str_replace(sampleid, '_short.+$','')) %>% 
  full_join(pheno, by = 'sampleid') %>% 
  ggboxplot(x = 'Toxicity', y = 'RPKM', color = 'Toxicity', palette = 'lancet', add = 'dotplot',
            title = 'CART samples secondary bile acid')  +
  stat_compare_means(comparisons = list(c('Y', 'N')), 
                     label = "p.signif",
                     method = 'wilcox.test',
                     correct = FALSE) 
```