### 2020-01-31 scripts/35_Braycurtis_distance_two_genera.Rmd
update

---
title: "Bray curtis distance to the group centroid"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(vegan)
library(pdist)
```

# entero_matrix

```{r}
sum_0_sample <- names(apply(entero_matrix, 2, sum)[apply(entero_matrix, 2, sum) == 0])
entero_matrix_fil <- entero_matrix[,!colnames(entero_matrix) %in% sum_0_sample]

# the N matrix and Y matrix
N_mat <- entero_matrix_fil[,colnames(entero_matrix_fil) %in% N_samps]
Y_mat <- entero_matrix_fil[,!colnames(entero_matrix_fil) %in% N_samps]

N_dist <- vegdist(t(N_mat), "bray")
Y_dist <- vegdist(t(Y_mat), "bray")

#dists <- pdist(t(N_mat), t(Y_mat))

all <- bind_rows(
    data_frame(group = 'N',distance = N_dist),
    data_frame(group = 'Y',distance = Y_dist)
    )

all %>% 
  ggboxplot(x = 'group', y = 'distance', add = 'jitter', title = 'entero_matrix') +
  stat_compare_means(
    comparisons = list(c('Y','N')),
    label = "p.signif",
    method = 'wilcox.test')
```

# bact_matrix

```{r}
sum_0_sample <- names(apply(bact_matrix, 2, sum)[apply(bact_matrix, 2, sum) == 0])

bact_matrix_fil <- bact_matrix[,!colnames(bact_matrix) %in% sum_0_sample]

# the N matrix and Y matrix
N_mat <- bact_matrix_fil[,colnames(bact_matrix_fil) %in% N_samps]
Y_mat <- bact_matrix_fil[,!colnames(bact_matrix_fil) %in% N_samps]

N_dist <- vegdist(t(N_mat), "bray")
Y_dist <- vegdist(t(Y_mat), "bray")

all <- bind_rows(
    data_frame(group = 'N',distance = N_dist),
    data_frame(group = 'Y',distance = Y_dist)
    )

all %>% 
  ggboxplot(x = 'group', y = 'distance', add = 'jitter', title = 'bact_matrix') +
  stat_compare_means(
    comparisons = list(c('Y','N')),
    label = "p.signif",
    method = 'wilcox.test')
```