### 2020-02-05 scripts/36_two_heatmap.Rmd
update

---
title: "Enterococcus and Bacteroidetes"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pheatmap)
library(tidyverse)
library(LaplacesDemon)
```

Investigating the contribution to the pathway abundances from the Bacteroidetes phylum and Enterococcus genus

```{r}
# load the normalized CPM count matrix with the stratified contribution from the above listed pylum and genus

entero_matrix <- read_tsv('../data/two_genera/genus_entero_pw_df.tsv') %>% 
    column_to_rownames('pw') %>% 
    as.matrix

bact_matrix <-  read_tsv('../data/two_genera/genus_bact_pw_df.tsv') %>% 
    column_to_rownames('pathway') %>% 
    as.matrix

pheno <- read_csv('../data/cleaned_pheno/final_pheno.csv')
```

## Visualize normalized counts in Bacteroidetes phylum

```{r out.width='200%', out.height='200%'}
annot <- data.frame(row.names = pheno$sampleid, tox = pheno$Toxicity, CR = pheno$CR)

pheatmap(
  bact_matrix,
  #color=bwrPalette, 
  annotation_col = annot,
  show_rownames = F,
  show_colnames = TRUE,
  #filename = '../figs/heatmap/bact_heatmap.jpg',
  height = 18,
  width = 28,
  cluster_rows = T,
  cluster_cols = T,
  main = 'All samples in Bacteroidetes'
)
```

Since 2098A seems to have too high counts and are making other samples counts ignorable. Try removing that one

```{r out.width='200%', out.height='200%'}
pheatmap(
  bact_matrix[,colnames(bact_matrix) != "2098A"],
  #color=bwrPalette, 
  annotation_col = annot,
  show_rownames = F,
  show_colnames = TRUE,
  #filename = '../figs/heatmap/bact_heatmap.jpg',
  height = 18,
  width = 28,
  cluster_rows = T,
  cluster_cols = T,
  main = 'Excluding sample 2098A in Bacteroidetes'
) 
```

It doesn't suggest any association between counts of certain pathways with the phenotype eg. toxicity.

## Binarize Enterococcus counts and visualize

Since the Enterococcus is only a genus and the counts are more sparse for that, binarize the matrix to presense -- 1, absense -- 0

```{r}
entero_matrix[entero_matrix > 0] = 1

pheatmap(
  entero_matrix,
  #color=bwrPalette, 
  annotation_col = annot,
  show_rownames = F,
  show_colnames = TRUE,
  #filename = '../figs/heatmap/entero_heatmap_binary.jpg',
  height = 18,
  width = 20,
  cluster_rows = T,
  cluster_cols = T,
  main = 'entero genus'
)
```

Again no differentiating between the groups

## Pairwise KL divergence

```{r}
library(arrangements)

combi <- combinations(colnames(bact_matrix), 2) %>% 
  as.data.frame(stringsAsFactors = F)

KLD_res <- apply(combi, 1, function(Row) {

  samp1_dens = density(bact_matrix[,Row[['V1']]])$y
  samp2_dens = density(bact_matrix[,Row[['V2']]])$y

  return(pairwise_KLD = KLD(samp1_dens, samp2_dens)$mean.sum.KLD	)

}) %>% 
  as.data.frame() %>% 
  bind_cols(combi) %>% 
  rename(KLD = names(.)[1])

KLD_res_matrix <- KLD_res %>% 
  spread(key = 'V2', value = 'KLD') %>% 
  column_to_rownames('V1') %>% 
  as.matrix()

pheatmap(
  KLD_res_matrix,
  #color=bwrPalette, 
  annotation_col = annot,
  show_rownames = T,
  show_colnames = T,
  #filename = '../figs/heatmap/entero_heatmap_binary.jpg',
  height = 18,
  width = 18,
  cluster_rows = T,
  cluster_cols = T,
  main = 'pairwise KL divergence'
)

```