### 2020-11-10 scripts/37_clean_pheno_refreshed.Rmd
update

---
title: "cleaning the pheno table"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

## use Antonio's script to clean the 16s data meta info

### msk

This table has the CR_d100

```{r}
#Oct/16/2019
#
#Loading patient info for Melody's project.
#
#
library('openxlsx');
library(data.table);
if(!exists("dt_carT_patient_info_with_sample_and_oligoID_time_stamp")){

  #Load patient info
  # dt_carT_patient_info = read.xlsx("/Volumes/castoricenter/CAR-Tcell_and_Microbiota/CTC Event Table 10162019.xlsx",
  #                                  sheetName = "Summary Table with Some Oligos");
  # dt_carT_patient_info = data.table(dt_carT_patient_info);
  # dt_carT_patient_info = dt_carT_patient_info[!is.na(pt_id)];
  # 
  #Convert sampleid to oligo
  dt_carT_sampleid_info = openxlsx::read.xlsx("/Volumes/castoricenter/CAR-Tcell_and_Microbiota/CTC Event Table01132020.xlsx",#CTC Event Table 11252019 wo Censor .xlsx",
                                              sheet = "Sample IDs only ");
                                    # sheetName = "Sample IDs only ");
  dt_carT_sampleid_info = data.table(dt_carT_sampleid_info);

  dt_carT_sampleid_info2 = openxlsx::read.xlsx("/Volumes/castoricenter/CAR-Tcell_and_Microbiota/CTC Event Table01132020.xlsx",#CTC Event Table 11252019 wo Censor .xlsx",
                                              sheet = "Summary Table with Some Oligos");
  dt_carT_sampleid_info2 = data.table(dt_carT_sampleid_info2);
  pt_id_2_mrn = dt_carT_sampleid_info2[!duplicated(pt_id)]$MRN
  names(pt_id_2_mrn) = dt_carT_sampleid_info2[!duplicated(pt_id)]$pt_id

  dt_carT_sampleid_info[,c("mrn_str","mrn"):=.(pt_id_2_mrn[pt_id], as.numeric(pt_id_2_mrn[pt_id])) ]

  #dt_carT_sampleid_info = dt_carT_sampleid_info[!is.na(p)]

  dt_carT_sampleid_info_m = melt.data.table(dt_carT_sampleid_info,id.vars = c("pt_id"),
                                            measure.vars = c("Baseline.Sample.ID",
                                                             "Week.1.Sample.ID",
                                                             "Week.2.Sample.ID",
                                                             "Week.3.Sample.ID",
                                                             "Week.4.Sample.ID"),
                                            variable.name = "time_group",
                                            value.name = "sampleid",
                                            na.rm = T);

  #sample ids with no time assignment:
  #d_set_tsne2[!sampleid %in% dt_carT_sampleid_info_m$sampleid ]$sampleid

    dt_carT_patient_info_with_sample_and_oligoID = merge(dt_carT_sampleid_info[is.na(Censor_reason), #Censor_YN=="N",
                                                                            .(pt_id,
                                                                              mrn,
                                                                              mrn_str,
                                                                              CR=`CR.(Y/N)`,
                                                                              ORR=`ORR.(Y/N)`,
                                                                              Toxicity=`Toxicity.(Y/N)`,
                                                                              #CR=CR..Y.N.,
                                                                              #ORR=ORR..Y.N..,
                                                                              #Toxicity=Toxicity..Y.N.,
                                                                              Cytokine.Release.Syndrome=`Cytokine.Release.Syndrome.(Y/N)`,
                                                                              #Cytokine.Release.Syndrome..Y.N..,
                                                                              ICANS_neurotoxicity = `ICANS.(Neurotoxicity)(Y/N)`,
                                                                              #ICANS_neurotoxicity = ICANS..Neurotoxicity..Y.N..,
                                                                              Disease.Class,
                                                                              Disease,
                                                                              antibiotic_prior=`Antibiotics.4.weeks.prior.to.CAR.T.cells.(Y/N)`,
                                                                              antibiotic_post=`Antibiotics.4.weeks.post-CAR.T.cells.(Y/N)`,
                                                                              Patient.Disease.Status.at.CAR,
                                                                              CAR.Group)],
                                                       dt_carT_sampleid_info_m[,.(pt_id,
                                                                                  sampleid,
                                                                                  time_group)])

  dt_carT_patient_info_with_sample_and_oligoID[, CR_or_ORR := factor(ifelse(CR=="Y" | ORR=="Y","Y","N"),levels=c("Y","N"))]

  dt_carT_patient_info_with_sample_and_oligoID[,
                                               time_group_simple :=factor(gsub("Sample.ID","",time_group))]

#   dt_carT_patient_info_with_sample_and_oligoID_time_stamp=Sys.time();
#   cat("patient info loaded in variable:\n\tdt_carT_patient_info_with_sample_and_oligoID\n")
# }else{
#   if(as.numeric(difftime(Sys.time(), dt_carT_patient_info_with_sample_and_oligoID_time_stamp, units="min")) > 1 ){
#     rm("dt_carT_patient_info_with_sample_and_oligoID_time_stamp");
#     source('~/projects/human_microbiota/library/Melody/load_carT_patient_info.R');
#   }else{
#     cat("patient info already loaded in variable:\n\tdt_carT_patient_info_with_sample_and_oligoID\n") 
#   }
  cat("patient info loaded in variable:\n\tdt_carT_patient_info_with_sample_and_oligoID\n")

}
```

### upenn

```{r}
#Nov/04/2020
#
#Load UPenn. CART patients clinical data for patients with stool samples.
#
#

if(!exists("dt_cart_stool_upen_samples_and_outcome")){

  dt_cart_stool_upen_clinical = openxlsx::read.xlsx("/Volumes/castoricenter/CAR-Tcell_and_Microbiota/U Penn/Microbiota DB_commercial CART_to send.xlsx",
                                                    sheet= 1)

  dt_cart_stool_upen_clinical = data.table(dt_cart_stool_upen_clinical);

  dt_cart_stool_upen_samples = xlsx::read.xlsx("/Volumes/castoricenter/CAR-Tcell_and_Microbiota/U Penn/Project 11198 Sample Key.xlsx",
                                               sheetIndex = 1);
  dt_cart_stool_upen_samples = data.table(dt_cart_stool_upen_samples)
  dt_cart_stool_upen_samples = dt_cart_stool_upen_samples[!is.na(Subject.ID)]
  dt_cart_stool_upen_samples$NA. = NULL;
  dt_cart_stool_upen_samples$NA..1 = NULL;

  dt_cart_stool_upen_samples_and_outcome = merge(dt_cart_stool_upen_samples[,.(sampleid=OligoID,pt_id=Subject.ID)],
                                                 dt_cart_stool_upen_clinical[,.(pt_id=`pt_id.(de-identified)`,
                                                                                CR=toupper(`CR(Y/N)`),
                                                                                Tox=`Toxicity(Y/N)`)])

}

print("Upen clincal+stool set is loaded in dt_cart_stool_upen_samples_and_outcome")

```

### italy

```{r}
#Oct/14/2020
#
#This scripts load Italian data, organizes it in antibiotics and clinical side. 
#
#

if(!(exists("dt_CART_italy_antibiotics") & exists("dt_CART_italy_clinical") & exists("dt_CART_italy_samples_and_outcome"))){

  dt_CART_italy = xlsx::read.xlsx("/Volumes/castoricenter/CAR-Tcell_and_Microbiota/Italy/DATABASE CAR T 091020.xlsx",
                                 sheetIndex = 1);
  dt_CART_italy = data.table(dt_CART_italy);
  setnames(dt_CART_italy,"pt_id.de.identified.","pid");
  dt_CART_italy = dt_CART_italy[!is.na(pid)]
  dt_CART_italy[, CR := ifelse(toupper(CR.Y.N.)=="Y",1,
                               ifelse(toupper(CR.Y.N.)=="N",0,
                               NA))]
  dt_CART_italy[, Tox := ifelse(toupper(Toxicity.Y.N.)=="Y",1,
                               ifelse(toupper(Toxicity.Y.N.)=="N",0,
                                      NA))]

  dt_CART_italy_clinical = dt_CART_italy[,infusion_date:=Data.infusione..dd.mm.yy.]

  dt_CART_italy_antibiotics = fread("/Volumes/castoricenter/CAR-Tcell_and_Microbiota/Italy/abx_data_john10-19-2020.csv")
  setnames(dt_CART_italy_antibiotics,"P_ID","pid");

  dt_CART_italy_antibiotics[, drug_name := tolower(drug_name)]
  dt_CART_italy_antibiotics[, drug_name := gsub("-","_",drug_name)]
  dt_CART_italy_antibiotics[, drug_name := gsub("/","_",drug_name)]

  dt_CART_italy_antibiotics[, start_date:=as.Date(start_date,"%m/%d/%y")]
  dt_CART_italy_antibiotics[, stop_date:=as.Date(stop_date,"%m/%d/%y")]

  # x=sort(tolower(unique(dt_CART_italy_antibiotics$drug_name)))
  # x=gsub("-","_",x)
  # x=unique(gsub("/","_",x))

  drug_name_clean_map = c("amoxicilina_ac.clavulanico"="amoxicillin",
                          "amoxicillin clavulanate"="amoxicillin",
                          "amoxicillin clavulanate _"="amoxicillin",
                          "augmentin"="amoxicillin",
                          "bactrim"="sulfamethoxazole_trimethoprim",
                          "cefixima"="cefixime",
                          "cefixoral"="cefixime",
                          "ceftazidima_avibactam"="ceftazidime",
                          "ceftriaxone vii somministrazione al"="ceftriaxone",
                          "clarithromycin"="clarithromycin",
                          "fosfomicina"="Fosfomycin",
                          "levofloxacina"="levofloxacin",
                          "teicoplanin"="teicoplanin",
                          "tigeciclina"="tigecycline",
                          "trimethoprim_sulfamethoxazole non_sop prophylaxis since diagnosis"="sulfamethoxazole_trimethoprim",
                          "vancomicina"= "vancomycin.pending")

  for(i in 1:length(drug_name_clean_map)){
    drug_name_cur = names(drug_name_clean_map)[i];
    drug_name_clean_cur = drug_name_clean_map[i];

    dt_CART_italy_antibiotics[drug_name %in% drug_name_cur, drug_name := drug_name_clean_cur];
  }

  #Adding sampleid and pid link here

  dt_CART_italy_samples = fread("/Volumes/castoricenter/CAR-Tcell_and_Microbiota/Italy/Set 81 Samples.txt")
  setnames(dt_CART_italy_samples,"pt_id","pid")
  setnames(dt_CART_italy_samples,"SampleID","sampleid")
  setnames(dt_CART_italy_samples,"OligoID","oligoid")

  dt_CART_italy_samples_and_outcome = merge(dt_CART_italy_samples[,.(pid,sampleid,oligoid,timepoint)],
                                            dt_CART_italy_clinical[,.(pid,CR,Tox)],
                                            all.x = T,
                                            by="pid")
  dt_CART_italy_samples_and_outcome[, CR:=.(ifelse(CR==1,"Y","N"))]
  dt_CART_italy_samples_and_outcome[, Tox:=.(ifelse(Tox==1,"Y","N"))]

}

print("dt_CART_italy_clinical, dt_CART_italy_antibiotics and dt_CART_italy_samples_and_outcome are loaded!")

```

## only look at the baseline samples for now (and we also need to have the CR out)

```{r}
# the msk table 
msk <- dt_carT_patient_info_with_sample_and_oligoID %>% 
  filter(time_group_simple == 'Baseline.')

upenn <- dt_cart_stool_upen_samples_and_outcome %>% 
  filter(!str_detect(sampleid, 'blank')) %>%  # two failed to amplify
  filter(!is.na(CR)) # one has not reached the CR record point

italy <- dt_CART_italy_samples_and_outcome %>% 
  filter(timepoint == 'PRE INFUSION') %>% 
  filter(!is.na(CR))  
```

CR:
day 30: msk and italy
day 100: upenn

Tox:
day 30: all of them

## getting the asv counts of those samples

```{r}
source('~/db_connect_simple.R')
connect_database(config_file = '~/dbConfig.txt')

get_table_from_database_predefined_filter('asv_counts_ag')
get_table_from_database('asv_sequences_ag')
```

```{r}
msk_cts <- asv_counts_ag %>% 
  filter(sampleid %in% msk$sampleid) %>% 
  select(asv_key, sampleid, count) %>% 
  spread(key = 'sampleid' , value = 'count', fill = 0)

# italy only has 3 in the db
italy_cts <- asv_counts_ag %>% 
  filter(sampleid %in% italy$oligoid) %>% 
  select(asv_key, sampleid, count) %>% 
  spread(key = 'sampleid' , value = 'count', fill = 0)

upenn_cts <- asv_counts_ag %>% 
  filter(sampleid %in% upenn$sampleid) %>% 
  select(asv_key, sampleid, count) %>% 
  spread(key = 'sampleid' , value = 'count', fill = 0)

# update italy pheno table
italy <- italy %>% 
  filter(oligoid %in% asv_counts_ag$sampleid)

```

## write out the asv counts and seqs for picrust2

```{r}
# write out the asv seq and cts to corresponding files
# ITS NOT FILTERED BUT TO ALIGN WITH THE SNAKEFILE FILE NAMES
library(seqinr)
cts_fns <- list(msk = msk_cts,
                italy = italy_cts,
                upenn = upenn_cts)

cts_fns %>% 
  imap(function(.x, .y){

    # the cts for the asv
    .x %>% 
      write_tsv(str_glue('../data/picrust/input/filtered_counts_{.y}.tsv'))

    # the seqs for the asv
    seqs = asv_sequences_ag %>%
        filter(asv_key %in% .x$asv_key)

    write.fasta(sequences = as.list(seqs$asv_sequence),
              names = seqs$asv_key,
              file.out = str_glue('../data/picrust/input/filtered_counts_asv_sequence_{.y}.tsv'))
  })
```

## load the downloaded results from picrust2

```{r}
src <- tibble(
  id = c('upenn','msk','italy'),
  local_dir = str_glue('/Users/daia1/projects/CART_microbiome/data/picrust/output/{id}'),
  pw_file = str_glue('daia1@lilac.mskcc.org:~/my_workdir/other_pipeline/picrust2/output/picrust2_out_{id}/pathways_out/path_abun_unstrat.tsv.gz'),
  dl_pw = str_glue('scp {pw_file} {local_dir}'),
  unzip_pw = str_glue('gunzip /Users/daia1/projects/CART_microbiome/data/picrust/output/{id}/path_abun_unstrat.tsv.gz')
)
# create local dir first
src %>% 
  mutate(mkdir_cmd = str_glue('mkdir {local_dir}')) %>% 
  select(mkdir_cmd) %>% 
  write_tsv('../data/mkdir.sh', col_names = F)

src %>% 
  select(dl_pw,  unzip_pw) %>% 
  gather('fn', 'cmd') %>% 
  select(cmd) %>% 
  write_tsv('../data/downloding_pw.sh', col_names = F)
```

## comparison together and individually

Lump italy and msk together since there are only three baseline samples from italy, and also italy and msk observe CR at day 30.

```{r}
pw_msk <- read_tsv('../data/picrust/output/msk/path_abun_unstrat.tsv')
pw_italy <- read_tsv('../data/picrust/output/italy/path_abun_unstrat.tsv')
pw_upenn <- read_tsv('../data/picrust/output/upenn/path_abun_unstrat.tsv')

# msk and italy
pheno_mitaly <- bind_rows(
  msk %>% 
    select(sampleid, CR, Toxicity),
  italy %>% 
    select(sampleid = oligoid,
           CR,
           Toxicity = Tox)
)

pw_mitaly <- pw_msk %>% 
  inner_join(pw_italy)

pw_mitaly_or <- bind_cols(pw_mitaly[,1],pw_mitaly[,pheno_mitaly$sampleid])

all.equal(colnames(pw_mitaly_or)[2:ncol(pw_mitaly_or)], pheno_mitaly$sampleid)

# upenn
pheno_upenn <- upenn %>% 
  mutate(sampleid = as.character(sampleid)) %>% 
  select(sampleid, CR, Toxicity = Tox)

pw_upenn_or <- bind_cols(pw_upenn[,1],pw_upenn[,pheno_upenn$sampleid])

all.equal(colnames(pw_upenn_or)[2:ncol(pw_upenn_or)], pheno_upenn$sampleid)

# all of them
pheno_all <- bind_rows(
  pheno_mitaly,
  pheno_upenn
)

pw_all <- pw_mitaly %>% 
  inner_join(pw_upenn)

pw_all_or <- bind_cols(pw_all[,1],pw_all[,pheno_all$sampleid])

all.equal(colnames(pw_all_or)[2:ncol(pw_all_or)], pheno_all$sampleid)
```
```{r}
pheno <- list(mitaly = pheno_mitaly,
              upenn = pheno_upenn,
              all = pheno_all)

pheno %>% 
  imap(~ select(.data = .x, CR) %>% 
            t %>% 
            write.table(str_glue('../data/lefse/{.y}__pheno_CR.txt'), sep = '\t', quote = F, row.names = T, col.names = F))

pheno %>% 
  imap(~ select(.data = .x, Toxicity) %>% 
            t %>% 
            write.table(str_glue('../data/lefse/{.y}__pheno_Toxicity.txt'), sep = '\t', quote = F, row.names = T, col.names = F))

cts_or <- list(
              mitaly = pw_mitaly_or,
              upenn = pw_upenn_or,
              all = pw_all_or
)

cts_or %>% 
  imap(~ .x %>% write_tsv(str_glue('../data/lefse/{.y}_cts.tsv')))

```

```{r}
# a script to bind those things and produce the correct format for the galaxy portal
cts <- list.files('../data/lefse/', pattern = 'cts')
pheno <- list.files('../data/lefse/', pattern = 'pheno')
combi <- expand.grid(cts_ = cts,pheno_ = pheno) %>% 
  mutate(cts_id = str_replace(cts_, '_cts.tsv','')) %>% 
  mutate(pheno_id = str_replace(pheno_, '__.+$','')) %>% 
  filter(cts_id == pheno_id) %>% 
  mutate(grp = if_else(str_detect(pheno_, 'CR'), 'CR', 'tox')) %>% 
  mutate(cmd = str_glue('cat {pheno_} {cts_} > {pheno_id}_{grp}.tsv'))

combi %>% select(cmd) %>% 
  write_tsv('../data/lefse/cat_table.sh', col_names = F)

```