### 2020-11-12 scripts/38_link_shotgun_fid_with_pheno.Rmd
jesus how much time I spent on this

---
title: "link the shotgun fid and the actual sample and outcome table"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

Actually the most important info with the sample: timepoint (so I know whether it is baseline) and the CR , Tox

```{r}
# the full italy samples and outcome joined from Antonio's two tables
italy_samp_outcome <- dt_CART_italy_samples %>% 
  full_join(dt_CART_italy_samples_and_outcome) %>% 
  distinct()

italy_samp_outcome %>% 
  write_csv('../data/shotgun_input/shotgun_italy_samp_outcome.csv')

# italy shotgun files
italy_fn <- read_csv('../../Catalog/data/full_project_11308_files.csv') %>% 
  distinct(sample_folder, .keep_all = T)

# there are 4 dups, and need to be concatenated to run
```

```{r}
samp_italy <- italy_fn %>% 
  left_join(italy_samp_outcome %>% mutate(fid = str_glue('{pid}{timepoint}')))

found1 <- samp_italy %>% 
  filter(!is.na(sampleid))

found2_df <- samp_italy %>% 
  filter(is.na(sampleid)) %>% 
  select(directory:fid) %>% 
  left_join(  italy_samp_outcome %>% 
    mutate(fid = str_replace_all(sampleid_raw, '\\+| |\\(|\\)|\\*','')  )
) %>% 
  distinct(fid, .keep_all = T) # there are two 16s samples CMT006 d172 and d168 both record as 180*

found2 <- found2_df %>% 
  filter(!is.na(sampleid))

found3_df <- found2_df %>% 
  filter(is.na(sampleid)) %>% 
  select(directory:fid) %>% 
  left_join(italy_samp_outcome %>% 
  filter(pid == 'CMT006') %>% 
  mutate(fid = str_replace_all(sampleid_raw, '\\+|\\(|\\)|\\*','')) %>% 
  filter(timepoint %in% c(90, 14))  %>% 
  mutate(fid = str_replace(fid, ' ','-'))) %>% 
  # there is one sample CMT006 that missed the datepoint that is actually day -3
  mutate(timepoint = if_else(fid == 'CMT006', -3.0, as.double(timepoint)),
         CR = if_else(fid == 'CMT006', 'Y', CR),
         Tox = if_else(fid == 'CMT006', 'N', Tox))

```

```{r}
# finally all of them matched ...

italy_pheno <- bind_rows(
  found1,
  found2,
  found3_df
) %>% 
  select(-sample_folder)

italy_pheno %>% 
  write_csv('../data/shotgun_pheno/italy_pheno_earlier.csv')
```

There are sevreral NA in the CR and tox in the baseline samples

Can you please clarify what “-B” when it is noted on the patient ID (pid)?
The “-B” for Patient ID CMT001B (sample ID “CMT001 PRE-BMT”) refers to a subsequent infusion for patient CMT001. Patient ID CMT001 after 42 days from his CAR T infusion received a BMT and it was recorded as Patient ID CMT001B and sampleID CMT001 PRE-BMT . Do not include Patient ID CMT001B and sampleID CMT001 PRE-BMT in the analysis

Can you please note once more the patients that should be excluded?
Exclude the patients in sponsored trial: CMT009; CMT011N; CMT013-B
Exclude the patients CMT001B for the reason at point “1”

so I think, CMT009; and ; CMT013-B ignore
what about CMT012 and CMT004B