### 2020-12-02 scripts/39_new_MSK_shotgun.Rmd
udpate

---
title: "New MSK samples"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

```{r}
# check the sequencing files returned, how the fid is matched to sampleid, is there any duplicates etc
proj <- c('Project_11454') # there are three failed QC so this is the correct number

folder <- '/Volumes/vandenBrinkLab/deep_sequencing'

files <- folder %>% 
  map_dfr(function(f){
    proj %>% 
      set_names(proj) %>% 
      imap_dfr( ~ Sys.glob(str_glue('{f}/{..1}/*/*')) %>% 
             as.data.frame %>% 
              mutate(..2)) %>% 
      rename(directory = names(.)[1],
             projectID = names(.)[2])
  }) %>% 
  mutate(sample_folder = str_extract(directory, 'Sample_.+$'),
         fid = str_replace(sample_folder, 'Sample_',''),
         fid = str_replace(fid, '_IGO_.+$','')) %>% 
  mutate(sampleid = fid) # the fid looks clean should be the sampleid

files %>% 
  distinct(sampleid) %>% 
  nrow
```

```{r}
# check the sample patient pheno table organized by John
# the first sheet is the eligible samples
stb <- readxl::read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/MSK/samples/sample_table_msk.xlsx', sheet = 1)

length(intersect(stb$sampleid, files$sampleid))

files %>% 
  write_csv('../data/shotgun_input/Project_11454_shotgun_samples.csv')
```