### 2020-12-08 scripts/41_new_I_E_shotgun.Rmd
update

---
title: "NEW Italy and Upenn samples"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

```{r}
source('~/db_connect_simple.R')
connect_database(config_file = '~/dbConfig.txt')
get_table_from_database('shotgun_lookup_ad')
```

## Italy ones

```{r}
ineligible <- readxl::read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/Italy/samples/opbg_sampleskey_with_oligoid_v2.xlsx', sheet = 2)

fntb_I <- shotgun_lookup_ad %>% 
  filter(projectid == 'Project_11453' & str_detect(fid, "^CMT")) %>% 
  # first see if that patient is eligible or not 
  mutate(pid = str_extract(fid, 'CMT...')) %>% 
  filter(! pid %in% ineligible$pt_id)  %>% 
  filter(!str_detect(fid, '_$')) %>% # the dups
  separate(fid, into = c('pre','day'), remove = F, convert = T, sep = '_') %>% 
  filter(is.na(day)) %>% # remove the ones I'm sure to be after infusion
  filter(fid != 'CMT029G0') %>%  # CMT029G0 is actually day 1 sample
  select(-pre, -day)
```
```{r}
fntb_P <- shotgun_lookup_ad %>%
  filter(projectid == 'Project_11453' & !str_detect(fid, "^CMT"))
```