### 2020-12-12 scripts/40_combine_all_samp_pheno.Rmd
good Im making progess and continue sorting things out

---
title: "combine all centers pheno"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

```{r}
# all centers pheno file 
# to get the baseline sampleid 
I <- read_csv('../data/shotgun_pheno/italy_all_earlier.csv') %>% 
  filter(timepoint < 0) %>% 
  ungroup() %>% 
  select(fid)

I2 <- read_csv('../data/shotgun_pheno/italy_all_later.csv') %>% 
  select(fid)

P <- read_csv('../data/shotgun_pheno/upenn_all_earlier.csv') %>% select(fid)

P2 <- read_csv('../data/shotgun_pheno/upenn_all_later.csv') %>% 
  select(fid)

M <- read_csv('../data/shotgun_pheno/msk_all_pheno.csv') %>% select(fid)

all_baseline <- bind_rows(I,I2, M, P, P2)

all_baseline %>% 
  write_csv('../data/shotgun_input/samples_cart_all_msk_patial_italy_upenn.csv', col_names = F)
```

## the pheno label for the samples

```{r}
I1 <- read_csv('../data/shotgun_pheno/italy_all_earlier.csv') %>% 
  filter(timepoint < 0) %>% 
  ungroup() %>% 
  select(fid, cr_d30:toxicity)

# the pheno lable for the I2!
iptb <- readxl::read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/Italy/patients/italy_essential_data_v4.xlsx', sheet = 1)
I2_pheno <- I2 %>% 
  rename(pt_id = pid) %>% 
  left_join(iptb) %>% 
  select(fid, cr_d30:toxicity)

I_phe <- bind_rows(
  I1, I2_pheno
) %>% 
  mutate(center = 'I')

# upenn
P_ptb <- readxl::read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/U Penn/patients/upenn_essential_data_v3.xlsx', sheet = 1)

P_ptb_ineligible <- readxl::read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/U Penn/patients/upenn_essential_data_v3.xlsx', sheet = 2)

P_ <- bind_rows(
  P %>% 
    mutate(pt_id = str_replace(fid, 'MS_UPENN_','')) %>% select(fid, pt_id) %>% mutate(pt_id = as.numeric(pt_id)),
  P2 %>% 
    mutate(pt_id = as.numeric(fid)) %>% select(fid, pt_id)
) %>% 
  filter(! pt_id %in% P_ptb_ineligible$pt_id)

P_phe <- P_ %>% 
  left_join(P_ptb)  %>% 
  select(fid, cr_d30:toxicity) %>% 
  select(fid, cr_d30, cr_d100, cr_overall, toxicity)  %>% 
  mutate(center = 'P')

# MSK
M_phe <- M %>% 
  select(fid, cr_d30:toxicity) %>% 
  mutate(center = 'M')

# combined

all <- bind_rows(
  M_phe, P_phe, I_phe
)

# a tally table
all %>% 
  count(center, toxicity)

all %>% 
  count(center, cr_d30)
```