### 2020-12-12 scripts/42_humann3.Rmd

---
title: "humann2 VS humann3"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggpubr)
```

```{r}
hm3 <- list.files('../data/log/', full.names = T, pattern = '_humann3.log$')

hm3_metric <- hm3 %>% 
  set_names(hm3) %>% 
  purrr::map(~ read_table(., col_names = F, col_types = 'c') %>% 
  rename(content = names(.)[1]) %>% 
  filter(str_detect(content, 'INFO: Unaligned')) %>% 
  transmute(hm_perc = str_extract(content, 'after.+$')) %>% 
  mutate(hm_perc = str_replace(hm_perc, 'after ','')) %>% 
  separate(hm_perc, into = c('type','perc'), sep = ': ') %>% 
  mutate(type = str_replace(type, ' ','_'),
         perc = as.numeric(str_replace(perc, ' %',''))) %>% 
  spread(key = 'type', value = 'perc')) %>% 
  bind_rows(.id = 'sampleid') %>% 
  # filter out the rows that have NA in translated column
  filter(!is.na(translated_alignment)) %>% 
  arrange(nucleotide_alignment) %>% 
  mutate(grp = 'hm3')

```

```{r}
hm2 <- hm3_metric %>% 
  mutate(sampleid = str_replace(sampleid, 'humann3','humann2')) %>% 
  pull(sampleid)

hm2_metric <- hm2 %>% 
  set_names(hm2) %>% 
  purrr::map(~ read_table(., col_names = F, col_types = 'c') %>% 
  rename(content = names(.)[1]) %>% 
  filter(str_detect(content, 'INFO: Unaligned')) %>% 
  transmute(hm_perc = str_extract(content, 'after.+$')) %>% 
  mutate(hm_perc = str_replace(hm_perc, 'after ','')) %>% 
  separate(hm_perc, into = c('type','perc'), sep = ': ') %>% 
  mutate(type = str_replace(type, ' ','_'),
         perc = as.numeric(str_replace(perc, ' %',''))) %>% 
  spread(key = 'type', value = 'perc')) %>% 
  bind_rows(.id = 'sampleid') %>% 
  # filter out the rows that have NA in translated column
  filter(!is.na(translated_alignment)) %>% 
  arrange(nucleotide_alignment) %>% 
  mutate(grp = 'hm2')
```

```{r}
all <- bind_rows(hm2_metric, hm3_metric) %>% 
  gather('type', 'value', nucleotide_alignment:translated_alignment) %>% 
  mutate(samples = str_replace(sampleid, '_humann.+$','')) %>% 
  mutate(samples = as.factor(samples)) %>% 
  mutate(perc_aligned = 100-value)

all %>% 
  ggboxplot(x = 'grp', y = 'perc_aligned', color = 'grp', facet.by = 'type', add = 'jitter', line.color = "gray", palette = 'lancet')

```