### 2020-12-17 scripts/45_ggplot_venn_lefse.Rmd
update

---
title: "ggplot and the venn diagram for the lefse bars !"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```