### 2021-02-15 scripts/45_16s_sorting_out.Rmd
update

---
title: "Sorting out the 16s"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(readxl)
```

This script will be cleaning and organizing the 16s samples

```{r}
# upenn is all baseline samples 
ustb <- read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/U Penn/samples/sample_mapping_UPenn.xlsx')

uptb <- read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/U Penn/patients/upenn_essential_data_v3.xlsx', sheet = 3) %>% 
  filter(!is.na(gender))

```

```{r}
# MSK contains baseline and following time points
mstb_ineli <- read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/MSK/samples/sample_table_msk.xlsx', sheet = 2)
mstb <- read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/MSK/samples/sample_table_msk.xlsx', sheet = 1) %>% 
  filter(! sampleid %in% mstb_ineli$sampleid)

mptb_ineli <-  read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/MSK/patients/msk_essential_data_v4.xlsx', sheet = 2)
mptb <-  read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/MSK/patients/msk_essential_data_v4.xlsx', sheet = 1) %>% 
  mutate(mrn = as.numeric(mrn)) %>% 
  filter(! mrn %in% mptb_ineli$pt_id)

```

## baseline samples only meta

```{r}
m_meta_b <- mstb %>% 
  filter(str_detect(time_group, '^Baseline')) %>% 
  inner_join(mptb, by = "mrn")

u_meta_b <- ustb %>% 
  inner_join(uptb, by = "pt_id")

colnames(u_meta_b)
```

```{r}
colnames(m_meta_b)
```

## Look at the tally of the phenotypes

```{r}
m_meta_b %>% 
  count(cr_d30)

m_meta_b %>% 
  count(cr_d100)

m_meta_b %>% 
  count(toxicity)
```
```{r}
u_meta_b %>% 
  count(cr_d30)

u_meta_b %>% 
  count(cr_d100)

u_meta_b %>% 
  count(toxicity)
```