### 2021-02-15 scripts/46_16s_relative_abundance.Rmd

---
title: "16s relative abundance"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(vdbR)
connect_database('~/dbConfig.txt')
library(tidyverse)
```

```{r}
get_table_from_database('asv_counts_ag')
get_table_from_database('asv_annotation_blast_ag')
```

```{r}
# the counts for the genus in MSK
mcts <- asv_counts_ag %>% 
  filter(sampleid %in% m_meta_b$sampleid) %>% 
  select(asv_key, sampleid, count) %>% 
  spread(key = 'sampleid', value = 'count', fill = 0) %>% 
  arrange(asv_key)

mannot <- asv_annotation_blast_ag %>% 
  filter(asv_key %in% mcts$asv_key) %>% 
  mutate(ordr =  if_else(ordr == '', str_glue('of_class_{class}'), ordr),
         family =  if_else(family == '', str_glue('of_order_{ordr}'), family),
         genus =  if_else(genus == '', str_glue('of_family_{family}'), genus),
         species =  if_else(species == '', str_glue('of_genus_{genus}'), species))

mcts_all <- mcts %>% 
  full_join(mannot %>%  select(asv_key, genus), by  = 'asv_key') %>% 
  select(-asv_key) %>% 
  gather(key = 'sampleid', value = 'count', names(.)[1]:names(.)[ncol(.) - 1]) %>% 
  group_by(sampleid, genus) %>% 
  summarise(cnt = sum(count)) %>% 
  # get the total count from the db to calculate the relab
  left_join(asv_counts_ag %>% distinct(sampleid,count_total ), by = 'sampleid') %>% 
  mutate(relab = cnt/count_total) %>% 
  select(sampleid, genus, relab) 
```
```{r}
# the counts for the genus in Upenn
ucts <- asv_counts_ag %>% 
  filter(sampleid %in% u_meta_b$oligoid) %>% 
  select(asv_key, sampleid, count) %>% 
  spread(key = 'sampleid', value = 'count', fill = 0) %>% 
  arrange(asv_key)

uannot <- asv_annotation_blast_ag %>% 
  filter(asv_key %in% ucts$asv_key) %>% 
  mutate(ordr =  if_else(ordr == '', str_glue('of_class_{class}'), ordr),
         family =  if_else(family == '', str_glue('of_order_{ordr}'), family),
         genus =  if_else(genus == '', str_glue('of_family_{family}'), genus),
         species =  if_else(species == '', str_glue('of_genus_{genus}'), species))

ucts_all <- ucts %>% 
  full_join(uannot %>%  select(asv_key, genus), by  = 'asv_key') %>% 
  select(-asv_key) %>% 
  gather(key = 'sampleid', value = 'count', names(.)[1]:names(.)[ncol(.) - 1]) %>% 
  group_by(sampleid, genus) %>% 
  summarise(cnt = sum(count)) %>% 
  # get the total count from the db to calculate the relab
  left_join(asv_counts_ag %>% distinct(sampleid,count_total ), by = 'sampleid') %>% 
  mutate(relab = cnt/count_total) %>% 
  select(sampleid, genus, relab) 
```
```{r}
# combine them and save to a df
genus <- bind_rows(
  ucts_all %>% mutate(center = 'P'),
  mcts_all %>% mutate(center = 'M')
)

genus %>% 
  write_csv('../data/amplicon/stool/counts_genus_combined.csv')
```