### 2021-02-22 scripts/49_16s_stacked_barplot.Rmd
udpate

---
title: "16s taxa stacked bar plot"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(vdbR)
connect_database('~/dbConfig.txt')
get_table_from_database("asv_annotation_blast_color_ag");
```

## load the tables and organize the data

```{r}
# my table of the CART stool cohort
stb <- read_csv('../data/amplicon/stool/combined_2_meta.csv')

# get the counts from database and also the color for the asv 
counts_data <- get_counts_subset(stb$sampleid)

counts_data %>% 
  distinct(sampleid)

dat <- counts_data %>% 
  select(asv_key:count_total, count_relative) %>% 
  left_join(asv_annotation_blast_color_ag %>% 
              select(asv_key,color_label_group_distinct), by = "asv_key")

# the color palette (inherited from Ying, used in lots of project in our lab, the palette used in the NEJM paper Fig 2D https://www.nejm.org/doi/full/10.1056/NEJMoa1900623)

asv_color_set <- asv_annotation_blast_color_ag %>% 
  distinct(color,color_label_group_distinct,color_label_group,color_base) %>% 
  select(color_label_group_distinct, color) %>% 
  deframe()
```

```{r}
# calculate the beta diversity between the samples which deicide the order of the samples in the plot
cbd <- compute_beta_diversity_and_tsne(sampleid = dat$sampleid, 
                                      taxonomy = dat$color_label_group_distinct,
                                      count = dat$count);
#compute beta diversity
cbd$compute_beta_diversity()
```

```{r}
#get beta diversity
d_beta <- cbd$get_betadiversity()

#compute hierarchical cluster
hc <- hclust(as.dist(d_beta), method = 'complete')

dend <- as.dendrogram(hc)

sample_dendogram_order <- labels(dend)

dat$sampleid = factor(dat$sampleid,levels = sample_dendogram_order)

ggplot(dat,aes(sampleid, count_relative, fill = color_label_group_distinct) ) +
  geom_bar(stat = "identity", position="fill", width = 1) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90),
        axis.text.y = element_blank(),
        legend.position = "none") +
  scale_fill_manual(values = asv_color_set)

```