### 2021-03-02 scripts/50_16s_diversity.Rmd
did this centroid thing anyway

---
title: "16s diversity"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggpubr)
```

```{r}
stb <- read_csv('../data/amplicon/stool/combined_2_meta.csv')
```
```{r}
# combining the institutions for the diversity boxplot colored by the toxicity and CR
dat <- stb %>% 
  select(sampleid, cr_d100, toxicity, simpson_reciprocal) %>% 
  gather('metric', 'value', cr_d100:toxicity)

# for coloring issues the two metric has to be done separately
dat %>% 
  filter(value %in% c('yes','no')) %>% 
  filter(metric == 'cr_d100') %>% 
  ggboxplot(x = 'value', y = 'simpson_reciprocal', add = 'jitter',
            title = 'cr_d100', ylab = 'Simpson Reciprocal', xlab = '', color = 'value',
            palette = c('#ED0000','#00468B')) +
            stat_compare_means(comparisons= list(c('yes', 'no')),
            label= "p.format",
            method= 'wilcox.test',
            correct=FALSE) +
  ggsave('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/Angel_amplicon/diversity/combined_cr_d100.pdf')

dat %>% 
  filter(value %in% c('yes','no')) %>% 
  filter(metric == 'toxicity') %>% 
  ggboxplot(x = 'value', y = 'simpson_reciprocal', add = 'jitter',
            title = 'toxicity', ylab = 'Simpson Reciprocal', xlab = '', color = 'value',
            palette = c('#0099B4','#AD002A'))  +
            stat_compare_means(comparisons= list(c('yes', 'no')),
            label= "p.format",
            method= 'wilcox.test',
            correct=FALSE) +
  ggsave('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/Angel_amplicon/diversity/combined_toxicity.pdf')

```