### 2021-03-02 scripts/52_shotgun_sig_pw_ontology.Rmd
this is actually pretty cool and it solves problems

---
title: "shotgun pathway ontology look"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

```{r}
pw <- read_csv('../data/shotgun_output/lefse_sig_results_taxa_and_pw.csv') %>% 
  filter(tbl == 'pathway') %>% 
  mutate(pwid = str_extract(feature, '^.+_PWY|^PWY_\\d{3,4}'))  %>% 
  mutate(pwid = str_replace_all(pwid, '_', '-'))
```

```{r}
level <- read_csv('~/pipeline/scripts/shotgun_pipeline/data/metacyc_pathway_class_and_superclass_levels.csv', col_types = 'ccccccccccccccccc')

dat <- pw %>% 
  inner_join(level)

dat_list <- dat %>% 
  arrange(cohort, outcome , direction, -score) %>% 
  split(list(.$cohort, .$outcome , .$direction)) %>% 
  discard( ~nrow(.) == 0)

dat_list %>% 
  imap(function(.x, .y){
    .x %>% 
      write_csv(str_glue('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/shotgun_results/lefse_sig_pathways_ontology_tables/{.y}_result.csv'))
  })
```