### 2021-03-08 scripts/48_16s_beta_diversity.Rmd

---
title: "16s beta diversity"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(vegan)
library(tidyverse)
set.seed(123)
```

Do a PcoA and do the anosim test

**Do M, P and pull separately, and then do for each of the 4 things CR_d100, Toxicity, ICANS and CRS**

```{r}
genus <- read_csv('../data/amplicon/stool/counts_genus_combined.csv')
meta <- read_csv('../data/amplicon/stool/combined_2_meta.csv') 
pheno <- meta %>% 
  select(sampleid, cr_d100, toxicity, crs, icans)

nsamp <- genus %>% distinct(sampleid) %>% nrow

# do some filtering 10% and 0.01%
keep <- genus %>% 
  filter(relab > 0.0001) %>% 
  count(genus) %>% 
  filter(n > floor(nsamp * 0.2)) %>% 
  filter(!is.na(genus)) %>% 
  pull(genus)

# find the dominating genus in each sample
domi <- genus %>% 
  arrange(sampleid, -relab) %>% 
  group_by(sampleid) %>% 
  slice(1) %>% 
  rename(domgenus = genus) %>% 
  select(sampleid, domgenus)

dat <- genus %>% 
  filter(genus %in% keep) %>% 
  spread(key = 'genus', value = 'relab') %>% 
  full_join(pheno) %>% 
  full_join(domi)

```

```{r}
# a function to do the combinations of the anosim test
pheno_ <- c(colnames(pheno)[2:5] , 'domgenus')
n <- length(pheno_)

combi <- data_frame(
  g1 = c(rep('M',n), rep('P',n), rep('M',n)),
  g2 = c(rep('M',n), rep('P',n), rep('P',n)),
  label = rep(pheno_, 3)
)

do_anosim_each <-  function(g1_, g2_, label_){
  g1_ = 'M'
  g2_ = 'P'
  label_ = 'domgenus'
    df = dat %>% 
      filter(center %in% c(g1_, g2_)) %>% 
      filter(get(label_) %in% c('yes','no')) 
    cts = df %>% 
      select(sampleid, all_of(keep)) %>% 
      column_to_rownames('sampleid')
    phe = df %>% 
      select(matches(label_)) %>% pull(get(label_)) 
    phedf = df %>% 
      select(sampleid, matches(label_))
    # 
    # # anosim run
    # ret = anosim(cts, phe, permutations = 999, distance = "bray", strata = NULL, parallel = 4)
    # 
    # # plot the pcoa
    dist_ <- vegdist(cts, method = 'bray')
    bc <- cmdscale(dist_, k = 2)
    # bc %>% 
    #   as.data.frame() %>% 
    #   rownames_to_column('sampleid') %>% 
    #   full_join(phedf, by = 'sampleid') %>% 
    #   ggplot(aes_string(x = "V1", y = "V2", color =  label_)) + geom_point() +
    #   labs(title = str_glue('PCoA {g1_}_{g2_}__{label_}\npval:{ret$signif} statistic:{round(ret$statistic, 2)}')) +
    #   ggsave(str_glue('../figs/beta_{g1_}_{g2_}__{label_}.pdf'))

    # the version without the anosim stats
    bc %>% 
      as.data.frame() %>% 
      rownames_to_column('sampleid') %>% 
      full_join(phedf, by = 'sampleid') %>% 
      ggplot(aes_string(x = "V1", y = "V2", color =  label_)) + geom_point() +
      labs(title = str_glue('PCoA {g1_}_{g2_}__{label_}')) +
      ggsave(str_glue('../figs/beta_{g1_}_{g2_}__{label_}.pdf'))

  # return(list(pval = ret$signif, 
  #             statistic = ret$statistic))
}

all_anosim <- pmap(combi, function(g1, g2, label){
    do_anosim_each(g1, g2, label)
  }) %>% 
  set_names(str_glue('{combi$g1}_{combi$g2}__{combi$label}')) 
```

```{r}
# to organize the final results
anosim_res <- all_anosim %>% 
  bind_rows(.id = 'name') %>% 
  separate(name, into = c('group','pheno'), remove = T, sep = '__') %>% 
  mutate(cohort = case_when(group == 'M_M' ~ "MSK", group == 'P_P' ~ "Upenn", group == 'M_P' ~ "Combined"))

anosim_res %>% 
  write_csv('../data/amplicon/stool/anosim_res.csv')
```

```{r}
# do the pcoa colored by the dominating genus 
do_domgenus_coloring_each <-  function(g1_, g2_, label_){
  g1_ = 'M'
  g2_ = 'P'
  label_ = 'domgenus'
    df = dat %>% 
      filter(center %in% c(g1_, g2_)) 
    cts = df %>% 
      select(sampleid, all_of(keep)) %>% 
      column_to_rownames('sampleid')
    phe = df %>% 
      select(matches(label_)) %>% pull(get(label_)) 
    phedf = df %>% 
      select(sampleid, matches(label_))

    # # plot the pcoa
    dist_ <- vegdist(cts, method = 'bray')
    bc <- cmdscale(dist_, k = 2)

    # the version without the anosim stats
    bc %>% 
      as.data.frame() %>% 
      rownames_to_column('sampleid') %>% 
      full_join(phedf, by = 'sampleid') %>% 
      ggplot(aes_string(x = "V1", y = "V2", color =  label_)) + geom_point() +
      labs(title = str_glue('PCoA  combined {label_}')) +
      ggsave(str_glue('../figs/beta_{g1_}_{g2_}__{label_}.pdf'))

}
```