### 2021-03-09 scripts/53_bayesian_5fam.Rmd
finally did some real coding after 5 PM

---
title: "Bayesian 5 taxa family"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggpubr)
library(brms)   
library(ggmcmc)
```

Model:
- CR_d100 ~ Faecalibacterium + Ruminococcus + Akkermansia + Bacteroides + Enterococcus

- Toxicity ~ Faecalibacterium + Ruminococcus + Akkermansia + Bacteroides + Enterococcus

```{r}
# the data
meta <- read_csv('../data/amplicon/stool/combined_2_meta.csv')
genus <- read_csv('../data/amplicon/stool/counts_genus_combined.csv') %>% 
  filter(genus %in% c('Faecalibacterium','Ruminococcus','Akkermansia','Bacteroides','Enterococcus')) %>% 
  select(-center) %>% 
  mutate(relablog = log10(relab + 2*10^-6)) %>% 
  select(-relab) %>% 
  spread(key = 'genus', value = 'relablog')

dat <- meta %>% 
  full_join(genus) %>% 
  mutate(toxicity = factor(toxicity, levels = c('no','yes')))

dat_cr <- dat %>% 
  filter(cr_d100 %in% c('yes','no')) %>% 
  mutate(cr_d100 = factor(cr_d100, levels = c('no','yes')))
```

```{r}
# the model
tox <- brm( toxicity  ~ 
              Akkermansia +
              Bacteroides +
              Enterococcus +
              Faecalibacterium +
              Ruminococcus +
               (1 | pt_id) ,  
              data = dat, 
              family = bernoulli(), 
              warmup = 1000, iter = 3000, 
              cores = 8, chains = 4, 
              seed = 123)
```