### 2021-03-09 scripts/53_bayesian_5fam.Rmd

---
title: "Bayesian 5 taxa family"
author: "Anqi Dai"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
```

```{r}
library(tidyverse)
library(ggpubr)
library(brms)   
library(ggmcmc)
```

Model:

- CR_d100 ~ Faecalibacterium + Ruminococcus + Akkermansia + Bacteroides + Enterococcus

- Toxicity ~ Faecalibacterium + Ruminococcus + Akkermansia + Bacteroides + Enterococcus

```{r}
# the data
meta <- read_csv('../data/amplicon/stool/combined_2_meta.csv')
genus <- read_csv('../data/amplicon/stool/counts_genus_combined.csv') %>% 
  filter(genus %in% c('Faecalibacterium','Ruminococcus','Akkermansia','Bacteroides','Enterococcus')) %>% 
  select(-center) %>% 
  mutate(relablog = log10(relab + 2*10^-6)) %>% 
  select(-relab) %>% 
  spread(key = 'genus', value = 'relablog')

dat <- meta %>% 
  full_join(genus) %>% 
  mutate(toxicity = factor(toxicity, levels = c('no','yes')))

dat_cr <- dat %>% 
  filter(cr_d100 %in% c('yes','no')) %>% 
  mutate(cr_d100 = factor(cr_d100, levels = c('no','yes')))
```

```{r}
# the model 
tox <- brm( toxicity  ~ 
              Akkermansia +
              Bacteroides +
              Enterococcus +
              Faecalibacterium +
              Ruminococcus +
               (1 | pt_id) ,  
              data = dat, 
              family = bernoulli(), 
              warmup = 5000, iter = 40000, 
              cores = 8, chains = 2, 
              seed = 123)

# It's giving errors: "Effective Samples Size (ESS) is too low, indicating posterior means and medians may be unreliable.
# Running the chains for more iterations may help"
```

```{r}
# Sean suggested to use individual genus first, good idea. 
# do 5 iteratively in a loop
target <- c('Akkermansia','Bacteroides','Enterococcus','Faecalibacterium','Ruminococcus')

ret <- target %>% 
  set_names(target) %>% 
  map(function(genus) {
      mod =  brm( as.formula(str_glue('toxicity  ~ 
              {genus} +
               (1 | pt_id)')),  
             data = dat, 
              family = bernoulli(), 
              warmup = 3000, iter = 10000, 
              cores = 8, chains = 2, 
              seed = 123)

      res = posterior_samples(mod) %>% 
              select(starts_with('b')) %>% 
              select(-b_Intercept)
      return(res)
  }) %>% 
  bind_cols()

```