### 2021-03-10 scripts/53_bayesian_5fam.Rmd

---
title: "Bayesian 5 taxa family"
author: "Anqi Dai"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
```

```{r}
library(tidyverse)
library(ggpubr)
library(brms)   
library(ggmcmc)
```

Model:

- CR_d100 ~ Faecalibacterium + Ruminococcus + Akkermansia + Bacteroides + Enterococcus

- Toxicity ~ Faecalibacterium + Ruminococcus + Akkermansia + Bacteroides + Enterococcus

```{r}
# the data
meta <- read_csv('../data/amplicon/stool/combined_2_meta.csv')
genus <- read_csv('../data/amplicon/stool/counts_genus_combined.csv') %>% 
  filter(genus %in% c('Faecalibacterium','Ruminococcus','Akkermansia','Bacteroides','Enterococcus')) %>% 
  select(-center) %>% 
  mutate(relablog = log10(relab + 2*10^-6)) %>% 
  select(-relab) %>% 
  spread(key = 'genus', value = 'relablog')

dat <- meta %>% 
  full_join(genus) %>% 
  mutate(toxicity = factor(toxicity, levels = c('no','yes')))

dat_cr <- dat %>% 
  filter(cr_d100 %in% c('yes','no')) %>% 
  mutate(cr_d100 = factor(cr_d100, levels = c('no','yes')))
```

```{r}
# the model 
# tox <- brm( toxicity  ~ 
#               Akkermansia +
#               Bacteroides +
#               Enterococcus +
#               Faecalibacterium +
#               Ruminococcus +
#                (1 | pt_id) ,  
#               data = dat, 
#               family = bernoulli(), 
#               warmup = 5000, iter = 40000, 
#               cores = 8, chains = 2, 
#               seed = 123)

# It's giving errors: "Effective Samples Size (ESS) is too low, indicating posterior means and medians may be unreliable.
# Running the chains for more iterations may help"
```

```{r}
# Sean suggested to use individual genus first, good idea. 
# do 5 iteratively in a loop
target <- c('Akkermansia','Bacteroides','Enterococcus','Faecalibacterium','Ruminococcus')

ret <- target %>% 
  set_names(target) %>% 
  map(function(genus) {
      mod =  brm( as.formula(str_glue('toxicity  ~ 
              {genus} +
               (1 | pt_id)')),  
             data = dat, 
              family = bernoulli(), 
              warmup = 3000, iter = 10000, 
              cores = 8, chains = 2, 
              seed = 123)

      res = posterior_samples(mod) %>% 
              select(starts_with('b')) %>% 
              select(-b_Intercept)
      return(res)
  }) %>%  
  bind_cols() %>% 
  gather('genus','coeff')
```

```{r}
# plotting in ridge plot
df <- ret %>% 
  group_by(genus) %>% 
  mutate(meanpergenus = mean(coeff))%>%
  ungroup()  %>% 
  mutate(genus = str_replace(genus, 'b_',''))

ggplot()+
  ggridges::geom_density_ridges(data  = df, 
                                aes(x      = coeff,
                                    y      = reorder(as.factor(genus), meanpergenus),
                                    height = ..density..),
                                scale = 1, size = 0.25,
                                alpha = 0.6) +
  geom_vline(xintercept = 0, col = "red", linetype = 'dashed', size = 1) +
  #scale_x_continuous(limits = c(-0.02,0.03)) +
  labs(y = '',
       title = 'Response: toxicity') +
  ggthemes::theme_tufte() +
  theme(axis.text.y = element_text(size=20),
        axis.text.x = element_text(size=15)) +
  ggsave('../figs/bayesian_genus_5_each_toxicity.jpg', width = 8)    
```

```{r}
# also when the response is cr_d100
ret_CR <- target %>% 
  set_names(target) %>% 
  map(function(genus) {
      mod =  brm( as.formula(str_glue('cr_d100  ~ 
              {genus} +
               (1 | pt_id)')),  
             data = dat_cr, 
              family = bernoulli(), 
              warmup = 3000, iter = 10000, 
              cores = 8, chains = 2, 
              seed = 123)

      res = posterior_samples(mod) %>% 
              select(starts_with('b')) %>% 
              select(-b_Intercept)
      return(res)
  }) %>%  
  bind_cols() %>% 
  gather('genus','coeff')
```
```{r}
ret_CR %>% 
  group_by(genus) %>% 
  mutate(meanpergenus = mean(coeff))%>%
  ungroup()  %>% 
  mutate(genus = str_replace(genus, 'b_','')) %>% 
  ggplot()+
  ggridges::geom_density_ridges(aes(x      = coeff,
                                    y      = reorder(as.factor(genus), meanpergenus),
                                    height = ..density..),
                                scale = 1, size = 0.25,
                                alpha = 0.6) +
  geom_vline(xintercept = 0, col = "red", linetype = 'dashed', size = 1) +
  #scale_x_continuous(limits = c(-0.02,0.03)) +
  labs(y = '',
       title = 'Response: cr_d100') +
  ggthemes::theme_tufte() +
  theme(axis.text.y = element_text(size=20),
        axis.text.x = element_text(size=15)) +
  ggsave('../figs/bayesian_genus_5_each_CR.jpg', width = 8)   
```