### 2021-03-19 scripts/58_shotgun_lefse_and_picked_target.Rmd
uypdate

---
title: "lefse and the picked targets"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggpubr)
```

```{r}
all_pw <- read_tsv('~/pipeline/scripts/shotgun_pipeline/data/metacyc_pathway_name_and_ID.tsv')
pick <- read_csv('../data/shotgun_output/metacyc_pathway_name_and_ID_MSv2.csv') %>% 
  select(Metabolite:cluster)

lefse <- read_csv('../data/shotgun_output/pull_pathway_lefse.csv') %>% 
  mutate(PWID = str_extract(feature, '^.+_PWY|^PWY.*_\\d{3,4}'))  %>% 
  mutate(
    PWID = case_when(
      str_detect(feature, 'NAD_BIOSYNTHESIS_II') ~ 'NAD_BIOSYNTHESIS_II',
      str_detect(feature, 'ARG_POLYAMINE_SYN') ~ 'ARG+POLYAMINE_SYN',
      str_detect(feature, 'PWY5F9') ~ 'PWY5F9',
      TRUE ~ PWID
    )
  ) %>% 
  mutate(PWID = str_replace_all(PWID, '_', '-'))

res <- lefse %>% 
  left_join(pick)

res %>% 
  write_csv('../data/shotgun_output/pull_lefse_results_pathway_join_with_picked_ones.csv')
```