### 2021-03-23 scripts/59_peruse_16s_lefse.Rmd
ud[ate

---
title: "Detailed look at the 16s lefse results"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggpubr)
```

```{r}
res <- list.files('../data/amplicon/lefse/', pattern = 'asv_tcts.tsv.res$', full.names = T)

res_all <- res %>% 
  set_names(res) %>% 
  map(~ read_tsv(., col_names = c('feature','xx','direction','score','pval')) %>% 
        filter(!is.na(score))) %>% 
  keep(~ nrow(.) > 0) %>% 
  bind_rows(.id = 'res') %>% 
  mutate(res = str_replace(res, '^.+//',''),
         res = str_replace(res, '_asv.+$',''))  %>% 
  rename(grp = res)
res_all$feature

```