### 2021-03-24 scripts/60_rethinking_g_clostridium.Rmd
update

---
title: "replacing bacteroides with Clostridium genus bayesian"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(rethinking)
library(tidyverse)  
library(ggpubr)
```

# CR/Toxicity ~ Akkermansia + Clostridium + Enterococcus + Faecalibacterium + Ruminococcus + center

## Toxicity

```{r}
# first I need to get the Clostridium genus abundance
clos <- read_csv('../data/amplicon/stool/counts_genus_combined.csv') %>% 
  filter(genus == 'Clostridium') %>% 
  mutate(relablog = log10(relab + 2*10^-6)) %>% 
  select(sampleid, Clostridium = relablog)

g5 <- read_csv('../data/amplicon/stool/combined_5_genera.csv') %>% 
  full_join(clos)

g5 %>% 
  write_csv('../data/amplicon/stool/combined_6_genera.csv')

meta <- read_csv('../data/amplicon/stool/combined_2_meta_expanded.csv')  %>% 
  full_join(g5)

set.seed(123)   
dat_list <- list(
    tox = meta$tox,
    cr100 = meta$cr100,
    location = meta$loca,
    Akkermansia = meta$Akkermansia,
    Clostridium = meta$Clostridium,
    Enterococcus = meta$Enterococcus,
    Faecalibacterium = meta$Faecalibacterium,
    Ruminococcus = meta$Ruminococcus
    ) 
```

```{r}
gtox <- ulam(
    alist(
        tox ~ dbinom( 1 , p ) ,
        logit(p) <- ba*Akkermansia + bc*Clostridium + be*Enterococcus + bf*Faecalibacterium + br*Ruminococcus + a[location] ,
        ba ~ dnorm( 0 , 1),
        bc ~ dnorm( 0 , 1),
        be ~ dnorm( 0 , 1),
        bf ~ dnorm( 0 , 1),
        br ~ dnorm( 0 , 1),
        a[location] ~ dnorm( 0 , 0.5 )
    ) , data=dat_list , chains=4 , log_lik=TRUE , cores = 8)
#precis(gtox, prob = 0.95)   
plot(precis(gtox, depth = 2, prob = 0.95) )  
```

```{r}
gcr <- ulam(
    alist(
        cr100 ~ dbinom( 1 , p ) ,
        logit(p) <- ba*Akkermansia + bc*Clostridium + be*Enterococcus + bf*Faecalibacterium + br*Ruminococcus + a[location] ,
        ba ~ dnorm( 0 , 1),
        bc ~ dnorm( 0 , 1),
        be ~ dnorm( 0 , 1),
        bf ~ dnorm( 0 , 1),
        br ~ dnorm( 0 , 1),
        a[location] ~ dnorm( 0 , 0.5 )
    ) , data=dat_list , chains=4 , log_lik=TRUE , cores = 8)
precis(gcr, prob = 0.95) 
plot(precis(gcr, depth = 2, prob = 0.95) ) 
```