### 2021-04-14 scripts/51_distance_2_healthy.Rmd

---
title: "16S distance to healthy people"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggpubr)
library(vdbR)
connect_database('~/dbConfig.txt')
source('centroid_distance.R')
get_table_from_database("asv_alpha_diversity_ag")
get_table_from_database("asv_annotation_blast_ag")
get_table_from_database("healthy_volunteers_ag")
stb <- read_csv('../data/amplicon/stool/combined_2_meta.csv')
```

```{r}
healthy <- healthy_volunteers_ag %>% 
  inner_join(asv_alpha_diversity_ag, by = c("sampleid", "oligos_id"))

cts <- get_counts_subset(c(stb$sampleid, healthy %>% pull(sampleid)))

# get the genus of the asv_key
dat <- cts %>% 
  left_join(asv_annotation_blast_ag %>% 
              select(asv_key, genus))

# calculate the beta diversity
cbd <- compute_beta_diversity_and_tsne(sampleid = dat$sampleid,
                                      taxonomy = dat$genus,
                                      count =  dat$count);

cbd$compute_beta_diversity()

d_beta = cbd$get_betadiversity() 
```

```{r}
# assemble the df with sampleid  mrn and group
df <- bind_rows(
  tibble(
    sampleid = stb$sampleid,
    pid = as.character(stb$pt_id),
    center = stb$center,
    grp = 'CART'
  ),
  tibble(
    sampleid = healthy$sampleid,
    pid = healthy$volunteer_uid,
    center = 'healthy',
    grp = 'healthy'
  )
)
```

```{r}
# compute the distance to centroid combined and for individual centers
centroid_grp <- centroid_distance_selected_rows(
  M = d_beta, 
  selected_rows = df$sampleid,
  selected_rows_groups = df$grp, 
  ref_group = "healthy")

centroid_center <- centroid_distance_selected_rows(
  M = d_beta, 
  selected_rows = df$sampleid,
  selected_rows_groups = df$center, 
  ref_group = "healthy")
```

```{r}
# plotting
centroid_grp %>% 
  ggboxplot(x = 'groups', y = 'dist_to_centroid', add = 'jitter', xlab = '',
            ylab = 'Distance to healthy cohort', title = 'Per sampleid')  +
  stat_compare_means(comparisons= list(c('CART', 'healthy')),
            label= "p.format",
            method= 'wilcox.test',
						correct=FALSE) +
  ggsave('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/Angel_amplicon/distance_to_centroid/all_CART_compared_to_healthy.pdf')
```
```{r}
centroid_center %>% 
  ggboxplot(x = 'groups', y = 'dist_to_centroid', add = 'jitter', xlab = '',
            ylab = 'Distance to healthy cohort', title = 'Per sampleid')  +
  stat_compare_means(comparisons= list(c('M','P'), c('M', 'healthy'), c('P', 'healthy')),
            label= "p.format",
            method= 'wilcox.test',
						correct=FALSE) +
  ggsave('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/Angel_amplicon/distance_to_centroid/centers_compared_to_healthy.pdf')

centroid_center %>% 
  filter(groups == 'healthy')
```
```{r}
# dominance in healthy volunteers 
h <- get_counts_subset(healthy$sampleid)

cts <- h %>% 
  select(asv_key, sampleid, count) %>% 
  spread(key = 'sampleid', value = 'count', fill = 0) %>% 
  arrange(asv_key)

mannot <- asv_annotation_blast_ag %>% 
  filter(asv_key %in% cts$asv_key) %>% 
  mutate(ordr =  if_else(ordr == '', str_glue('of_class_{class}'), ordr),
         family =  if_else(family == '', str_glue('of_order_{ordr}'), family),
         genus =  if_else(genus == '', str_glue('of_family_{family}'), genus),
         species =  if_else(species == '', str_glue('of_genus_{genus}'), species)) 
mcts_all <- cts %>% 
  full_join(mannot %>%  select(asv_key, genus), by  = 'asv_key') %>% 
  select(-asv_key) %>% 
  gather(key = 'sampleid', value = 'count', names(.)[1]:names(.)[ncol(.) - 1]) %>% 
  group_by(sampleid, genus) %>% 
  summarise(cnt = sum(count)) %>% 
  # get the total count from the db to calculate the relab
  left_join(h %>% distinct(sampleid,count_total ), by = 'sampleid') %>% 
  mutate(relab = cnt/count_total) %>% 
  select(sampleid, genus, relab)

mcts_all %>% 
  write_csv('../data/amplicon/healthy_volunteer_genus_relab.csv')

h %>% 
  write_csv('../data/amplicon/healthy_volunteer_asv_relab.csv')
```