### 2021-04-14 scripts/57_16s_lefse.Rmd
update

---
title: "16s lefse"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggpubr)
```

```{r}
# sort out the genera counts table and also do filtering (need to have all taxa levels)
meta <- read_csv('../data/amplicon/stool/combined_2_meta.csv')

library(vdbR)
connect_database('~/dbConfig.txt')
get_table_from_database('asv_annotation_blast_ag')
cts <- get_counts_subset(meta$sampleid)

cts_ <- cts %>% 
  select(asv_key, sampleid, count_relative) %>% 
  spread(key = 'sampleid', value = 'count_relative', fill = 0)

annot <- asv_annotation_blast_ag %>% 
  filter(asv_key %in% cts_$asv_key) %>% 
  mutate(ordr =  if_else(ordr == '', str_glue('of_class_{class}'), ordr),
         family =  if_else(family == '', str_glue('of_order_{ordr}'), family),
         genus =  if_else(genus == '', str_glue('of_family_{family}'), genus),
         species =  if_else(species == '', str_glue('of_genus_{genus}'), species)) %>% 
  mutate(taxa_asv = str_glue('k__{kingdom}|p__{phylum}|c__{class}|o__{ordr}|f__{family}|g__{genus}|s__{species}|a__{asv_key}'))

cts_all <- cts_ %>% 
  full_join(annot %>%  select(asv_key, taxa_asv), by  = 'asv_key') %>% 
  select(-asv_key)  %>% 
  gather('sampleid', 'relab', names(.)[1]:names(.)[ncol(.)-1])

# the asv to keep
keepg <- cts_all %>% 
  filter(relab > 0) %>%
  ungroup() %>% 
  count(taxa_asv) %>% 
  filter(n > floor(nrow(meta) * 0.1)) %>% 
  pull(taxa_asv)

cts_fil <- cts_all %>% 
  filter(taxa_asv %in% keepg) %>% 
  spread('sampleid', 'relab')
```

```{r}
# dat <- genera %>% 
#   filter(genus == 'Ruminococcus') %>% 
#   gather('sampleid', 'relab', names(.)[2]:names(.)[ncol(.)]) %>% 
#   full_join(meta) %>% 
#   mutate(logrelab =  log10(relab + 2e-6)) 
# 
# yes <- dat %>% 
#   filter(cr_d100 == 'yes')
# no <- dat %>% 
#   filter(cr_d100 == 'no')
# 
# 
# dat %>% 
#   #filter(center == 'M') %>% 
#   mutate(logrelab =  log10(relab + 2e-6)) %>% 
#   ggboxplot(x = 'cr_d100', y = 'logrelab', add = 'jitter') +
# stat_compare_means(comparisons= list(c('yes', 'no')),
# 										label= "p.signif",
# 										method= 'wilcox.test',
# 										correct=FALSE)
# 
# t.test(yes$logrelab, no$logrelab)

```

```{r}
# the pheno label for the samples
pheno <- meta %>% 
  select(center, cr_d100:crs, icans, sampleid) %>% 
  gather('pheno', 'value', cr_d100:icans)

all_sub_pheno <- pheno %>% 
  split(., list(.$pheno)) %>% 
  purrr::imap(~ filter(.data = ., value != 'not_assessed'))

tpheno <- all_sub_pheno %>% 
  imap(function(.x, .y){
    select(.data = .x, value) %>% 
      t() %>% write.table(str_glue('../data/amplicon/lefse/pull_{.y}.txt'), sep = '\t', quote = F, row.names = T, col.names = F)
  })

tcts <- all_sub_pheno %>% 
  map(~ pull(.data = ., sampleid) ) %>% 
  imap(~ cts_fil %>% select(taxa_asv, matches(.x)) %>%  write_tsv(str_glue('../data/amplicon/lefse/{.y}_asv_tcts.tsv'))) 
```

cat pull_toxicity.txt toxicity_asv_tcts.tsv > pull_toxicity_asv_tcts.tsv
cat pull_cr_d100.txt cr_d100_asv_tcts.tsv > pull_cr_d100_asv_tcts.tsv 
cat pull_crs.txt crs_asv_tcts.tsv > pull_crs_asv_tcts.tsv
cat pull_icans.txt icans_asv_tcts.tsv > pull_icans_asv_tcts.tsv

```{r}
fns <- list.files('../data/amplicon/lefse/', pattern = 'pull.*_asv_tcts.tsv$')

cmds <- tibble(
  fns = fns
) %>% 
  mutate(format_cmd = str_glue('format_input.py {fns}  {fns}.in -c 1  -u 2 -o 1000000   ')) %>% 
  mutate(run_cmd = str_glue('run_lefse.py {fns}.in  {fns}.res')) %>% 
  mutate(plot_cmd = str_glue('plot_res.py {fns}.res {fns}.pdf --format pdf  --feature_font_size 4 --width 10 --dpi 300 --title {fns}')) %>% 
  mutate(clado_cmd = str_glue('plot_cladogram.py {fns}.res {fns}_clado.pdf  --label_font_size 4 --dpi 300 --format pdf --title {fns}')) %>% 
  select(-fns) %>% 
  gather() %>% 
  select(value) %>% 
  write_csv('../data/amplicon/lefse/lefse_run_cmd_taxa.sh', col_names = F)

# No differentially abundant features found in the 4
# also this is the case when using center as subclass 
```