### 2021-07-01 scripts/63_CLR_transformation.Rmd

---
title: "CLR transformation"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(compositions)
library(tidyverse)
```

```{r}
genus <- read_csv('../data/amplicon/stool/counts_genus_combined.csv') %>% 
  select(-center) %>% 
  spread(key = 'sampleid', value = 'relab', fill = 0) %>% 
  mutate(genus = if_else(is.na(genus), 'NA',genus)) %>% 
  column_to_rownames('genus')

clr_res <- clr(genus) %>% 
  as.data.frame()  %>% 
  rownames_to_column('genus') %>% 
  gather('sampleid','clr', names(.)[2]:names(.)[ncol(.)])

clr_5 <- clr_res %>% 
  filter(genus %in% c('Faecalibacterium','Ruminococcus','Akkermansia','Bacteroides','Enterococcus'))

clr_5 %>% 
  write_csv('../data/amplicon/stool/counts_genus_combined_CLR.csv')
```