### 2021-07-13 scripts/62_logistic_regression_version.Rmd

---
title: "logistic regression"
author: "Anqi Dai"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F)
```

```{r}
library(magrittr) 
library(ggeffects)
library(tidyverse)
library(lme4)
library(cowplot)
library(grid)
library(ggpubr)
library(sjPlot)
theme_set(theme_sjplot())
```

```{r}
meta <- read_csv('../data/amplicon/stool/combined_2_meta.csv') %>% 
  mutate(center = factor(center),
         toxicity = factor(toxicity, levels = c('no','yes')),
         cr_d100 = factor(cr_d100, levels = c('no','yes')))

dat <- meta %>% 
  select(toxicity,cr_d100, simpson_reciprocal, center)

dat %>% 
  head
```

# Alpha diversity as predictor

## Toxicity ~ alpha diversity + center

**The ggeffects package computes estimated marginal means (predicted values) for the response, at the margin of specific values or levels from certain model terms, i.e. it generates predictions by a model by holding the non-focal variables constant and varying the focal variable(s). ggpredict() uses predict() for generating predictions**

```{r}
m_tox <- glm(
  toxicity ~ simpson_reciprocal + center, 
  data = dat, 
  family = binomial(link = "logit")
)

summary(m_tox)

#ggpredict(m_tox,se=TRUE,interactive=TRUE,digits=3, c('simpson_reciprocal')) %>% plot()

# plot_model(m_tox, transform = NULL) +
#   ggsave('../figs/amplicon/diversity_tox_logi.pdf', height = 3, width = 4)

or_CI <- round(cbind(coef(m_tox), confint(m_tox)), digits=3) %>% 
  as.data.frame()

or_CI <- or_CI %>% 
  mutate(variable=rownames(or_CI)) %>% 
  rename(lower_bound = `2.5 %` , upper_bound = `97.5 %`)

# Now plot them
or_CI[2,] %>%  #remove row number 1 (The intercept)
  ggplot(aes(x = reorder(variable, V1), y = V1)) +
  geom_point(shape = 15,
             size  = 4, width = 0.1,
             position = "dodge", color="black") +
  geom_errorbar(aes(ymin  = lower_bound,
                    ymax  = upper_bound),
                width = 0.1,
                size  = 0.7,
                position = "dodge", color="turquoise4") +
  theme(axis.title = element_text(face = "bold")) +
  xlab("Variables") + ylab("Coefficient with 95% CI") +
  coord_flip(ylim = c(-0.5, 1)) +
  #geom_hline(yintercept = 1, color = "red", size = 1) +
   theme(axis.title = element_text(size = 0)) +
  theme(axis.text = element_text(size = 14)) +
  labs(title = 'Toxicity') +
  ggsave('../figs/amplicon/logis_div_tox.pdf', width = 5, height = 4)

```

## CR ~ alpha diversity + center

```{r}
m_CR <- glm(
  cr_d100 ~ simpson_reciprocal + center, 
  data = dat, 
  family = binomial(link = "logit")
)

or_CI <- round(cbind(coef(m_CR), confint(m_CR)), digits=3) %>% 
  as.data.frame()

or_CI <- or_CI %>% 
  mutate(variable=rownames(or_CI)) %>% 
  rename(lower_bound = `2.5 %` , upper_bound = `97.5 %`)

#Now plot them
or_CI[2,] %>%  #remove row number 1 (The intercept)
  ggplot(aes(x = reorder(variable, V1), y = V1)) +
  geom_point(shape = 15,
             size  = 4, width = 0.1,
             position = "dodge", color="black") +
  geom_errorbar(aes(ymin  = lower_bound,
                    ymax  = upper_bound),
                width = 0.1,
                size  = 0.7,
                position = "dodge", color="turquoise4") +
  theme(axis.title = element_text(face = "bold")) +
  xlab("Variables") + ylab("Coefficient with 95% CI") +
  coord_flip(ylim = c(-0.5, 1)) +
  #geom_hline(yintercept = 1, color = "red", size = 1) +
   theme(axis.title = element_text(size = 0)) +
  theme(axis.text = element_text(size = 14)) +
  labs(title = 'CR') +
  ggsave('../figs/amplicon/logis_div_cr.pdf', width = 5, height = 4)
```
```{r}
# plot_model(m_CR, transform = NULL, terms = c('simpson_reciprocal')) +
#   ggsave('../figs/amplicon/diversity_cr_logi.pdf', height = 3, width = 4)
# 
# get_model_data(m_CR, type = "est")
# log(get_model_data(m_CR, type = "est")$estimate)
```

```{r}
#ggpredict(m_CR,se=TRUE,interactive=TRUE,digits=3, c('simpson_reciprocal')) %>% plot()
```

# 5 genera as predictor

## Toxicity ~ Akkermansia + Bacteroides + Enterococcus + Faecalibacterium + Ruminococcus + center

The genera's relative abundance is log10 transformed.

```{r}
genera <- read_csv('../data/amplicon/stool/combined_5_genera.csv')

# the CLR transformed genera abundance 
clr <- read_csv('../data/amplicon/stool/counts_genus_combined_CLR.csv')  %>% 
  spread(key = 'genus', value = 'clr')

gmeta <- meta  %>% 
  inner_join(clr) %>% 
  select(toxicity,cr_d100, Akkermansia,Bacteroides, Enterococcus, Faecalibacterium, Ruminococcus , center)

gmeta %>% 
  head
```

```{r}
g_tox <- glm(
  toxicity ~ Akkermansia + Bacteroides + Enterococcus + Faecalibacterium + Ruminococcus + center, 
  data = gmeta, 
  family = binomial(link = "logit")
)
summary(g_tox)

gb <- ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Bacteroides')) %>% plot()
gr <- ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Ruminococcus')) %>% plot()
ge <- ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Enterococcus')) %>% plot()
ga <- ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Akkermansia')) %>% plot()
gf <- ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Faecalibacterium')) %>% plot()

plot_model(g_tox)+
  ggsave('../figs/amplicon/genera_tox_logi_CLR.pdf', height = 3, width = 4)
```

```{r}
gtox_clr <- ggarrange(gb, gr, ge, ga, gf,
    ncol = 1
    #o = 'hv',
    #labels = 'AUTO',
    #axis = 'bltr'
    )  
```

## CR ~ Akkermansia + Bacteroides + Enterococcus + Faecalibacterium + Ruminococcus + center

```{r}
g_CR <- glm(
  cr_d100 ~ Akkermansia + Bacteroides + Enterococcus + Faecalibacterium + Ruminococcus + center, 
  data = gmeta, 
  family = binomial(link = "logit")
)
summary(g_CR)

gb <- ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Bacteroides')) %>% plot()
gr <- ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Ruminococcus')) %>% plot()
ge <- ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Enterococcus')) %>% plot()
ga <- ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Akkermansia')) %>% plot()
gf <- ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Faecalibacterium')) %>% plot()

plot_model(g_CR)+
  ggsave('../figs/amplicon/genera_cr_logi_CLR.pdf', height = 3, width = 4)
```

```{r}
gCR_clr <- ggarrange(gb, gr, ge, ga, gf,
    ncol = 1
    #o = 'hv',
    #labels = 'AUTO',
    #axis = 'bltr'
    ) +
    ggsave(str_glue('../figs/amplicon/gcr_CLR_5.pdf'), width = 3, height = 8, dpi = 300)
```

# log 10 transformed version of the genera relab

```{r}
gmeta_log10 <- meta  %>% 
  inner_join(genera) %>% 
  select(toxicity,cr_d100, Akkermansia,Bacteroides, Enterococcus, Faecalibacterium, Ruminococcus , center)

```

```{r}
g_tox <- glm(
  toxicity ~ Akkermansia + Bacteroides + Enterococcus + Faecalibacterium + Ruminococcus + center, 
  data = gmeta_log10, 
  family = binomial(link = "logit")
)
summary(g_tox)
gb <- ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Bacteroides')) %>% plot()
gr <- ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Ruminococcus')) %>% plot()
ge <- ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Enterococcus')) %>% plot()
ga <- ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Akkermansia')) %>% plot()
gf <- ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Faecalibacterium')) %>% plot()
gtox_log <- ggarrange(gb, gr, ge, ga, gf,
    ncol = 1
    #o = 'hv',
    #labels = 'AUTO',
    #axis = 'bltr'
    ) +
    ggsave(str_glue('../figs/amplicon/gtox_log10_5.pdf'), width = 3, height = 8, dpi = 300)

plot_model(g_tox) +
  ggsave('../figs/amplicon/genera_tox_logi_log10.pdf', height = 3, width = 4)
```

## CR ~ Akkermansia + Bacteroides + Enterococcus + Faecalibacterium + Ruminococcus + center

```{r}
g_CR <- glm(
  cr_d100 ~ Akkermansia + Bacteroides + Enterococcus + Faecalibacterium + Ruminococcus + center, 
  data = gmeta_log10, 
  family = binomial(link = "logit")
)
summary(g_CR)
gb <- ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Bacteroides')) %>% plot()
gr <- ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Ruminococcus')) %>% plot()
ge <- ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Enterococcus')) %>% plot()
ga <- ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Akkermansia')) %>% plot()
gf <- ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Faecalibacterium')) %>% plot()
gCR_log <- ggarrange(gb, gr, ge, ga, gf,
    ncol = 1
    #o = 'hv',
    #labels = 'AUTO',
    #axis = 'bltr'
    )

plot_model(g_CR)+
  ggsave('../figs/amplicon/genera_cr_logi_log10.pdf', height = 3, width = 4)
```

# side-by-side comparison of the CLR and log10 transformed

```{r}
tox_all <- ggarrange(gtox_log, gtox_clr, nrow = 1) +
    ggsave(str_glue('../figs/amplicon/gtox_all_5.pdf'), width = 6, height = 8, dpi = 300)

cr_all <- ggarrange(gCR_log, gCR_clr, nrow = 1) +
    ggsave(str_glue('../figs/amplicon/gCR_all_5.pdf'), width = 6, height = 8, dpi = 300)
```