### 2021-07-13 scripts/64_Bayesian_with_bifido.Rmd

---
title: "bayesian with Bifido"
author: "Anqi Dai"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F)
```

```{r}
library(rethinking)
library(tidyverse)  
library(ggpubr)
set.seed(123)
```

# CR/Toxicity ~ Akkermansia + Bacteroides + Enterococcus + Faecalibacterium + Ruminococcus + Bifidobacterium + center

```{r}
bifido <- read_csv('../data/amplicon/stool/counts_genus_combined.csv') %>% 
  filter(genus == 'Bifidobacterium') %>% 
  select(-center) %>% 
  spread(key = 'genus', value = 'relab') %>% 
  mutate(Bifidobacterium = log10(Bifidobacterium + 2*10^-6))

genera <- read_csv('../data/amplicon/stool/combined_5_genera.csv') %>% 
  inner_join(bifido)

# do not standardize the log transformed relative abundance 
meta <- read_csv('../data/amplicon/stool/combined_2_meta_expanded.csv')  %>% 
  inner_join(genera) %>% 
  mutate(diseaseID = if_else(disease == 'NHL', 1, 2))

# wanna see the range of the 5 genera log 10 relab
meta %>% select(Akkermansia:Bifidobacterium) %>% 
  summary
```

```{r}
set.seed(123)   
dat_list <- list(
    tox = meta$tox,
    cr100 = meta$cr100,
    location = meta$loca,
    Akkermansia = meta$Akkermansia,
    Bacteroides = meta$Bacteroides,
    Enterococcus = meta$Enterococcus,
    Faecalibacterium = meta$Faecalibacterium,
    Ruminococcus = meta$Ruminococcus,
    Bifidobacterium = meta$Bifidobacterium,
    did = meta$diseaseID
    ) 
```

```{r}
gtox <- ulam(
    alist(
        tox ~ dbinom( 1 , p ) ,
        logit(p) <- ba*Akkermansia + bb*Bacteroides + be*Enterococcus + bf*Faecalibacterium + br*Ruminococcus + bd*Bifidobacterium + a[location] ,
        ba ~ dnorm( 0 , 1),
        bb ~ dnorm( 0 , 1),
        be ~ dnorm( 0 , 1),
        bf ~ dnorm( 0 , 1),
        br ~ dnorm( 0 , 1),
        bd ~ dnorm( 0 , 1),
        a[location] ~ dnorm( 0 , 0.5 )
    ) , data=dat_list , chains=4 , log_lik=TRUE , cores = 8)
precis(gtox, prob = 0.95)   
plot(precis(gtox, depth = 2, prob = 0.95) )  
# the  Ruminococcus looks more significant now
```

```{r}
gcr <- ulam(
    alist(
        cr100 ~ dbinom( 1 , p ) ,
        logit(p) <- ba*Akkermansia + bb*Bacteroides + be*Enterococcus + bf*Faecalibacterium + br*Ruminococcus + bd*Bifidobacterium + a[location] ,
        ba ~ dnorm( 0 , 1),
        bb ~ dnorm( 0 , 1),
        be ~ dnorm( 0 , 1),
        bf ~ dnorm( 0 , 1),
        br ~ dnorm( 0 , 1),
        bd ~ dnorm( 0 , 1),
        a[location] ~ dnorm( 0 , 0.5 )
    ) , data=dat_list , chains=4 , log_lik=TRUE , cores = 8)
precis(gcr, prob = 0.95) 
plot(precis(gcr, depth = 2, prob = 0.95) ) 
# no posterior prediction 
# genera from literature , 
```

# add disease to the original model

```{r}
gtox_d <- ulam(
    alist(
        tox ~ dbinom( 1 , p ) ,
        logit(p) <- ba*Akkermansia + bb*Bacteroides + be*Enterococcus + bf*Faecalibacterium + br*Ruminococcus + d[did] + a[location] ,
        ba ~ dnorm( 0 , 1),
        bb ~ dnorm( 0 , 1),
        be ~ dnorm( 0 , 1),
        bf ~ dnorm( 0 , 1),
        br ~ dnorm( 0 , 1),
        d[did] ~ dnorm( 0 , 0.5 ),
        a[location] ~ dnorm( 0 , 0.5 )
    ) , data=dat_list , chains=4 , log_lik=TRUE , cores = 8)
precis(gtox_d, prob = 0.95)   
plot(precis(gtox_d, depth = 2, prob = 0.95) )  
# the disease effect is not strong as center 
```

```{r}
gcr_d <- ulam(
    alist(
        cr100 ~ dbinom( 1 , p ) ,
        logit(p) <- ba*Akkermansia + bb*Bacteroides + be*Enterococcus + bf*Faecalibacterium + br*Ruminococcus + d[did] + a[location] ,
        ba ~ dnorm( 0 , 1),
        bb ~ dnorm( 0 , 1),
        be ~ dnorm( 0 , 1),
        bf ~ dnorm( 0 , 1),
        br ~ dnorm( 0 , 1),
        d[did] ~ dnorm( 0 , 0.5 ),
        a[location] ~ dnorm( 0 , 0.5 )
    ) , data=dat_list , chains=4 , log_lik=TRUE , cores = 8)
precis(gcr_d, prob = 0.95) 
plot(precis(gcr_d, depth = 2, prob = 0.95) ) 
```

```{r}
# the distribution of the outcome in each of the disease
table(meta$disease, meta$cr_d100)
table(meta$disease, meta$toxicity)
```