### 2021-07-19 scripts/63_CLR_transformation.Rmd

---
title: "CLR transformation"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(compositions)
library(tidyverse)
library(ggpubr)
```
CLR works with counts, not relab. And also need to add a pseudocount

```{r}
meta <- read_csv('../data/amplicon/stool/combined_2_meta.csv')

library(vdbR)
connect_database('~/dbConfig.txt')
get_table_from_database('asv_annotation_blast_ag')
cts <- get_counts_subset(meta$sampleid)
cts_ <- cts %>% 
  select(asv_key, sampleid, count)

annot <- asv_annotation_blast_ag %>% 
  filter(asv_key %in% cts_$asv_key)

cts_all <- cts_ %>% 
  left_join(annot %>%  select(asv_key, genus), by  = 'asv_key') %>% 
  group_by(sampleid, genus) %>% 
  summarise(cnt = sum(count)) %>% 
  spread('sampleid', 'cnt', fill = 0) %>% 
  filter(!is.na(genus)) %>% 
  column_to_rownames('genus')
```

```{r}
clr_res <- clr(cts_all + 0.5) %>% 
  as.data.frame()  %>% 
  rownames_to_column('genus') %>% 
  gather('sampleid','clr', names(.)[2]:names(.)[ncol(.)])

clr_5 <- clr_res %>% 
  filter(genus %in% c('Faecalibacterium','Ruminococcus','Akkermansia','Bacteroides','Enterococcus'))

clr_5 %>% 
  write_csv('../data/amplicon/stool/counts_genus_combined_CLR.csv')
```

```{r}
log10 <- read_csv('../data/amplicon/stool/combined_5_genera.csv') %>% 
  gather('genus', 'log10', Akkermansia:Ruminococcus)

two <- log10 %>% 
  inner_join(clr_5)

two %>% 
  arrange(log10) %>% 
  #filter(log10 > -5) %>% 
  ggscatter(x = 'log10', y = 'clr', facet.by = 'genus', alpha = 0.4,
             add = "reg.line",  # Add regressin line
           add.params = list(color = "blue", fill = "lightgray"), # Customize line
           conf.int = TRUE, # Add confidence interval
           cor.coef = TRUE, # Add correlation coefficient.
           cor.coeff.args = list(method = "pearson",  label.sep = "\n") )
```