### 2021-07-28 scripts/69_lefse_shotgun_pathway.Rmd

---
title: "lefse shotgun pathway"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

```{r}
simple <- read_csv('../data/shotgun_pheno/final_comprehensive_UPDATED_simple.csv')
pheno <- simple %>% 
  gather('pheno', 'value', cr_d100:crs) 
```

```{r}
# the pathway counts table
full <- read_tsv('../data/shotgun_output/humann3_tsv/humann3_pathabundance_CART.tsv')  %>% 
  rename_all(funs(str_replace(., '^CART_','')))  %>% 
  rename_all(funs(str_replace(., '_humann3$','')))

all_sub_pheno <- pheno %>% 
  split(., list(.$pheno)) %>% 
  purrr::imap(~ filter(.data = ., value != 'not_assessed'))

all_sub_pheno %>% 
  imap(function(.x, .y){
    select(.data = .x, value) %>% 
      t() %>% 
      write.table(str_glue('../data/shotgun_output/humann3/pull_{.y}.txt'), sep = '\t', quote = F, row.names = T, col.names = F)
  })

all_pcts <- all_sub_pheno %>% 
  map(~ pull(.data = ., fid) ) %>% 
  imap(~ full %>% select(`# Pathway`, matches(.x)) %>% 
         write_tsv(str_glue('../data/shotgun_output/humann3/pull_{.y}_pcts.tsv')))
```

Snakemake do the normalization and the split

```{r}
# add a filtering step here 
# 50 and 25% 
pcts <- read_tsv('../data/shotgun_output/humann3/humann_final_out/pull_cr_d100_pcts_cpm_unstratified.tsv') %>% 
  gather('sampleid', 'cpm', names(.)[2]:names(.)[ncol(.)]) %>% 
  rename(pw = `# Pathway`)

keeppw <- pcts %>% 
  filter(cpm > 50) %>%
  ungroup() %>% 
  count(pw) %>% 
  filter(n > floor(nrow(simple) * 0.25)) %>% 
  pull(pw)

cts_fil <- pcts %>% 
  filter(pw %in% keeppw) %>% 
  spread('sampleid', 'cpm', fill = 0)

cts_fil %>% 
  write_tsv('../data/shotgun_output/humann3/humann_final_out/pull_cr_d100_pcts_cpm_unstratified_fil.tsv')
```

```{r}
# tox
pcts <- read_tsv('../data/shotgun_output/humann3/humann_final_out/pull_toxicity_pcts_cpm_unstratified.tsv') %>% 
  gather('sampleid', 'cpm', names(.)[2]:names(.)[ncol(.)]) %>% 
  rename(pw = `# Pathway`)

keeppw <- pcts %>% 
  filter(cpm > 50) %>%
  ungroup() %>% 
  count(pw) %>% 
  filter(n > floor(nrow(simple) * 0.25)) %>% 
  pull(pw)

cts_fil <- pcts %>% 
  filter(pw %in% keeppw) %>% 
  spread('sampleid', 'cpm', fill = 0)

cts_fil %>% 
  write_tsv('../data/shotgun_output/humann3/humann_final_out/pull_toxicity_pcts_cpm_unstratified_fil.tsv')
```

Snakemake do the binding

```{r}
# PATHWAY
# I already normalized the pcts so don't need to normalize again here 
fns <- list.files('../data/shotgun_output/humann3/', pattern = 'lefse_ready_pcts.tsv$')

cmds <- tibble(
  fns = fns
) %>% 
  mutate(format_cmd = str_glue('format_input.py {fns}  {fns}.in -c 1 -u 2')) %>% 
  mutate(run_cmd = str_glue('run_lefse.py {fns}.in  {fns}.res')) %>% 
  mutate(plot_cmd = str_glue('plot_res.py {fns}.res {fns}.pdf --format pdf  --feature_font_size 4 --width 10 --dpi 300 --title {fns}')) %>% 
  select(-fns) %>% 
  gather() %>% 
  select(value) %>% 
  write_csv('../data/shotgun_output/humann3/lefse_run_cmd.sh', col_names = F)
```

```{r}
# look at the pathway results
library(vdbR)
connect_database('~/dbConfig.txt')
get_table_from_database('metacyc_pathway_name')
get_table_from_database('metacyc_pathway_ontology')

fns <- list.files('../data/shotgun_output/humann3/', pattern = 'lefse_ready_pcts.tsv.res$', full.names = T)

feature <- fns %>% 
  set_names(fns) %>% 
  map(~ read_tsv(., col_names = c('pathway','xx','direction','score','pval')) %>% 
        filter(!is.na(score))) %>% 
  bind_rows(.id = 'group') %>% 
  mutate(group = str_replace(group, '../data/shotgun_output/humann3//','')) %>% 
  mutate(group = str_replace(group, '_lefse_ready_pcts.tsv.res$',''))

# change the "N" direction to be minus score
feature <- bind_rows(
  feature %>% 
    split(.$direction) %>% 
    pluck('no') %>% 
    mutate(score = -score),
  feature %>% 
    split(.$direction) %>% 
    pluck('yes')  
) %>% 
  arrange(group, pathway, score) %>% 
  mutate(pwid = str_extract(pathway, '^.+_PWY|^PWY.*_\\d{3,4}')) %>% 
  mutate(pwid = str_replace_all(pwid, '_', '-')) %>% 
  mutate(pwid = if_else(str_detect(pathway, '^TCA'), 'TCA', pwid)) %>% 
  mutate(pwid = if_else(str_detect(pathway, '^NAD'), 'NAD-BIOSYNTHESIS-II', pwid)) %>% 
  inner_join(metacyc_pathway_name %>% select(pwid, pw_name)) %>% 
  inner_join(metacyc_pathway_ontology %>% select(pwid, l4:l9))

all_title_fs <- 20
axis_text_fs <- 16
CR <- feature %>% 
  filter(group == 'pull_cr_d100') %>% 
  ggplot(aes(x = reorder(pathway, score), y = score, fill = direction)) +
            geom_bar( stat = 'identity') +
            coord_flip() +
            scale_color_manual(values = c('#925E9F', '#42B540'))  +
            scale_fill_manual(values = c('#925E9F', '#42B540')) +
            theme_classic() +
            theme(axis.title.y  = element_blank(), 
                  plot.title = element_text(size=all_title_fs),
                  axis.title.x = element_text(size=axis_text_fs),
                  axis.text.x = element_text(size=axis_text_fs),
                  legend.position='bottom') +
            labs(title = str_glue('Response to CR') ,
                 y = 'Score')

tox <- feature %>% 
  filter(group == 'pull_toxicity') %>% 
  ggplot(aes(x = reorder(pathway, score), y = score, fill = direction)) +
            geom_bar(stat = 'identity') +
            coord_flip() +
            scale_color_manual(values = c('#0099B4', '#AD002A'))  +
            scale_fill_manual(values = c('#0099B4', '#AD002A')) +
            theme_classic() +
            theme(axis.title.y  = element_blank(), 
                  axis.title.x = element_text(size=axis_text_fs),
                  plot.title = element_text(size=all_title_fs),
                  axis.text.x = element_text(size=axis_text_fs),
                  legend.position="bottom") +
            labs(title = str_glue('Response to Toxicity') ,
                 y = 'Score')

cowplot::plot_grid(CR,tox,  
          nrow = 2, 
          rel_heights = c(1,3),
          align = 'hv',
          axis = 'b')
```